---
title: "Read IFQuant data"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

In this script, we will read in the mIF data that was processed by IFQuant.

## Load data

We will read in the tsv files that contain cells from the same region as 
the IMC acquisition and store them in `SingleCellExperiment` objects.

```{r load-data, message=FALSE}
library(SingleCellExperiment)
library(tidyverse)

cur_files <- list.files("~/SIB Swiss Institute of Bioinformatics/Robin Liechti - Workflow2021/Lausanne_data/ROI_registration_20220523/",
                        recursive = TRUE, 
                        pattern = "cells_properties_ROI_coordinates.tsv.gz", full.names = TRUE)

cur_sces <- lapply(cur_files, function(x){
  cur_data <- read.csv(x, sep = "\t")
  cur_counts <- t(cur_data[,grepl("score.normalized", colnames(cur_data))])
  cur_raw <- t(cur_data[,grepl("score$", colnames(cur_data))])
  rownames(cur_counts) <- rownames(cur_raw) <- str_split(rownames(cur_raw), "\\.", 
                                                         simplify = TRUE)[,1]
  cur_sce <- SingleCellExperiment(assays = list(counts = cur_counts,
                                                raw = cur_raw))
  colData(cur_sce) <- DataFrame(cur_data[,!grepl("score", colnames(cur_data))])
  colData(cur_sce)$sample_id <- str_split(x, "/", simplify = TRUE)[,9]
  colData(cur_sce)$patient_id <- str_extract(cur_sce$sample_id, "[0-9]{8}")
  colData(cur_sce)$ROI <- str_split(cur_sce$sample_id, "_", simplify = TRUE)[,5]
  return(cur_sce)
})

cur_colnames <- Reduce(intersect, lapply(cur_sces, function(x){colnames(colData(x))}))
cur_sces <- lapply(cur_sces, function(x){
  colData(x) <- colData(x)[,cur_colnames]
  return(x)
})

sce <- do.call("cbind", cur_sces)

# Transform counts
assay(sce, "exprs") <- asinh(counts(sce))
```

## Cell type annotation

We will now annotate the cells based on the celltype annotation key.

```{r cell-type-annotation, message=FALSE}
cur_key <- as.data.frame(read_csv("~/switchdrive/Institution/IMMUCan/Imaging_working_group/data_for_sharing/workflow_samples/mIF/phenotype_key.csv"))

rownames(cur_key) <- cur_key$phenotype

sce$celltype <- cur_key[sce$phenotype,2]
sce$matched_celltype <- sce$celltype
sce$matched_celltype[grepl("Tumor", sce$matched_celltype)] <- "Tumor"
```

## Define colors 

Here, we define colors for matched cell types and samples.

```{r define-colors}
library(RColorBrewer)
color_vectors <- list()
matched_celltype <- setNames(c("#3F1B03", "#0A3A03", "#066970", "#894F36", "#636160", "#DC6FE2", "#BF0A3D", "#F4800C"),
                             c("Tumor", "Tcell", "Neutrophil", "Macrophage",
                               "other", "DC", "Bcell", "BnTcell"))
patient_id <- setNames(brewer.pal(8, "Set1"),
                       c("10061074", "10067433", "10068868", "10071582", "10073140", "10074349", "10074832", "10075371"))
color_vectors$matched_celltype <- matched_celltype
color_vectors$patient_id <- patient_id

metadata(sce)$color_vectors <- color_vectors
```

## Add interaction graph

Finally, we will add a 20µm expansion interaction graph.

```{r buildSpatialGraph, message=FALSE}
library(imcRtools)
sce <- buildSpatialGraph(sce, img_id = "sample_id", type = "expansion", 
                         coords = c("nucleus.x", "nucleus.y"),
                         threshold = 20, name = "expansion_20")
```

## Save object

Finally, we will save the generated object

```{r}
saveRDS(sce, "/Volumes/immucan_volume/processed_data/Panel_1/2022_WORKFLOW/mIF/Rout/sce_ROIs.rds")
```


