---
title: "Read IFQuant data"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# set path
the git repo contains a vector which specifies the path to the data folder. please adapt this path in your clone to run the repo automated.
```{r}
mount_path <- readRDS(file = "data/mount_path.rds")
```

In this script, we will read in the mIF data that was processed by IFQuant.

## Define color vectors

These will be shared across all datasets.

```{r define-colors}
library(RColorBrewer)

color_vectors <- list()
matched_celltype <- setNames(c("sienna4", "magenta3", "skyblue3", "green3", "grey75", "green4", "#BF0A3D", "#F4800C"),
                             c("Tumor", "T", "Neutrophil", "MacCD163",
                               "other", "DC", "B", "BnT"))
celltype <- setNames(c("sienna4", "tan3", "magenta3", "skyblue3", "green3", "grey75", "green4", "#BF0A3D", "#F4800C"),
                        c("Tumor", "Tumor_CD15", "T", "Neutrophil", "MacCD163",
                          "other", "DC", "B", "BnT"))

patient_id <- setNames(c("#BE1E2D", "#FF4B5B", "#F7941D", "#FFCA89", "#00A651", "#5FFF73", "#00AEEF", "#9BD7FF", "#2E3192", "#675BFF"),
                          c("10061074", "10067433", "10068868", "10071582", "10073140", "10074349", "10074832", "10075371", "10075572", "10082495"))

col_vec_indication <- c(CRC = "#5B1C55", SCCHN = "#39BEB4", NSCLC = "#F79C1D", 
                        BCC = "#3F85A7", RCC = "#C81F43")

color_vectors$patient_id <- patient_id
color_vectors$Indication <- col_vec_indication
color_vectors$matched_celltype <- matched_celltype
color_vectors$celltype <- celltype
```

# Read in the single-cell data

We use `imcRtools` to read in the data:

```{r read-data}
sce <- read_steinbock(paste0(mount_path,"steinbock_2025/"), 
                      return_as = "sce",
                      image_file = NULL)
sce$image <- paste0(sce$sample_id, ".tiff")
sce$patient_id <- sub("_", "", str_extract(sce$image, "_[0-9]{8}"))
assay(sce, "exprs") <- asinh(counts(sce))
head(colData(sce))

# Save color vectors in metadata slot
metadata(sce)$color_vectors <- color_vectors
```


```{r save-RDS, warning=FALSE}
cur_dir <- paste0(mount_path,"steinbock_2025/Rout/")
if (!dir.exists(cur_dir)) dir.create(cur_dir)
saveRDS(sce, paste0(mount_path,"steinbock_2025/Rout/sce.rds"))
```

# Read in image data

Here, we read in the images and masks and save them on disk.

```{r read-images, message = FALSE, warning=FALSE}
library(cytomapper)
images <- loadImages(paste0(mount_path,"steinbock_2025/img/"), pattern = "*.tiff")
masks <- loadImages(paste0(mount_path,"steinbock_2025/masks/"),
                    as.is = TRUE, pattern = "*.tiff")

channelNames(images) <- rownames(sce)

mcols(images) <- mcols(masks) <- DataFrame(image = names(images),
                                           sample_id = sub(".tiff", "", names(images)))

saveRDS(images, paste0(mount_path,"steinbock_2025/Rout/images.rds"))
saveRDS(masks, paste0(mount_path,"steinbock_2025/Rout/masks.rds"))
```

# Read in the nuclear single-cell data

We use `imcRtools` to read in the data:

```{r read-data}
sce_nuc <- read_steinbock(paste0(mount_path,"steinbock_2025/"), 
                      return_as = "sce",
                      image_file = NULL,
                      intensities_folder = "intensities_nucleus",
                      regionprops_folder = "regionprops_nucleus",
                      graphs_folder = "neighbors_nucleus")

sce_nuc$image <- paste0(sce_nuc$sample_id, ".tiff")
sce_nuc$patient_id <- sub("_", "", str_extract(sce_nuc$image, "_[0-9]{8}"))
assay(sce_nuc, "exprs") <- asinh(counts(sce_nuc))
head(colData(sce_nuc))

# Save color vectors in metadata slot
metadata(sce_nuc)$color_vectors <- color_vectors
```


```{r save-RDS, warning=FALSE}
cur_dir <- paste0(mount_path,"steinbock_2025/Rout/")
if (!dir.exists(cur_dir)) dir.create(cur_dir)
saveRDS(sce_nuc, paste0(mount_path,"steinbock_2025/Rout/sce_nuc.rds"))
```

# Read in image data

Here, we read in the images and masks and save them on disk.

```{r read-images, message = FALSE, warning=FALSE}
library(cytomapper)
masks <- loadImages(paste0(mount_path,"steinbock_2025/masks_nucleus/"),
                    as.is = TRUE, pattern = "*.tiff")

channelNames(images) <- rownames(sce)

mcols(images) <- mcols(masks) <- DataFrame(image = names(images),
                                           sample_id = sub(".tiff", "", names(images)))
saveRDS(masks, paste0(mount_path,"steinbock_2025/Rout/masks_nuc.rds"))
```
