---
title: "Sample effect corrections"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# set path
this path should be the directory where all files downloaded from Zenodo are stored in the same structure as on Zenodo
```{r}
mount_path <- "O:/projects/immucan/processed_data/Panel_1/2022_WORKFLOW/"
```

In the quality control section, we observe sample differences between tumor cells across the samples.
We can now correct for these effect using batch effect correction methods developed for scRNAseq data.

## Load data and libraries

First, we will load the data and libraries again.

```{r load-data-and-libraries, message=FALSE}
library(batchelor)
library(dittoSeq)
library(scater)
library(viridis)
library(DT)
library(BiocSingular)
library(cowplot)

sce <- readRDS(paste0(mount_path,"IMC/Rout/sce.rds"))
```

## Perform samples correction

Here, we will used the mutual nearest neighbour approach of the `batchelor`
package to integrate cells from different patients.

```{r fastMNN}
set.seed(220321)
out <- fastMNN(sce, batch = sce$patient_id,
               auto.merge = TRUE,
               subset.row = rowData(sce)$use_channel,
               assay.type = "exprs",
               d = 30,
               BSPARAM = ExactParam())

# Transfer the correction results to the main spe object
reducedDim(sce, "fastMNN") <- reducedDim(out, "corrected")
```

## Quality control of the correction results

We can now highlight samples that were more difficult to integrate than others.

```{r integration-QC}
merge_info <- metadata(out)$merge.info 
datatable(data.frame(DataFrame(left = merge_info$left,
          right = merge_info$right,
          batch.size = merge_info$batch.size,
          max_lost_var = rowMax(merge_info$lost.var))))
```

## Visualize corrected cells

As a visual assessment of sample correction, we visualize the integrated cells
as UMAP.

```{r dimred}
set.seed(220321)
sce <- runUMAP(sce, dimred = "fastMNN", name = "UMAP_fastMNN")
```

We can now visualize the patient IDs and expression on the UMAPs.

```{r dimred-viz-patient-indication, message=FALSE}
dittoDimPlot(sce, var = "patient_id", reduction.use = "UMAP_fastMNN", size = 0.2) + 
    scale_color_manual(values = metadata(sce)$color_vectors$patient_id) +
    ggtitle("Patient ID on UMAP")

dittoDimPlot(sce, var = "indication", reduction.use = "UMAP_fastMNN", size = 0.2) + 
    scale_color_manual(values = metadata(sce)$color_vectors$Indication) +
    ggtitle("Patient ID on UMAP")
```

```{r dimred-viz-expression, fig.height=12, fig.width=20}
p.list <- list()
for(i in rownames(sce)){
  p.list[[i]] <- plotReducedDim(sce, colour_by = i, by_exprs_values = "exprs", 
                          point_size = 0.1, point_alpha = 1, dimred = "UMAP_fastMNN") 
}

plot_grid(plotlist = p.list, ncol = 7)
```

## Save object

Finally, we will save the modified `SingleCellExperiment` object.

```{r save-RDS}
saveRDS(sce, paste0(mount_path,"IMC/Rout/sce.rds")
```
