---
title: "Quality control"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# set path
the git repo contains a vector which specifies the path to the data folder. please adapt this path in your clone to run the repo automated.
```{r}
mount_path <- readRDS(file = "data/mount_path.rds")
```


Here, we will perform quality control on all samples with focus
on batch effects, cell number per images and cell size.

## Load data and libraries

First, we will load all relevant data and libraries.

```{r load-data, message=FALSE}
library(tidyverse)
library(dittoSeq)
library(scater)
library(viridis)
library(DT)
library(cowplot)

sce <- readRDS(paste0(mount_path,"IMC/Rout/sce.rds"))
```

## Cell numbers

We will now visualize how many cells we detected per image and patient.

```{r cell-numbers}
# Plot number of cells
colData(sce) %>%
  as_tibble() %>%
  group_by(patient_id, ROI) %>%
  summarize(count = n()) %>%
  ggplot() +
    geom_bar(aes(patient_id, count, fill = ROI), 
             position = "stack", stat = "identity") +
    scale_fill_manual(values = metadata(sce)$color_vectors$ROIs) + 
    theme_minimal(base_size = 15) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Numeric output
colData(sce) %>%
  as_tibble() %>%
  group_by(patient_id, ROI) %>%
  summarize(count = n()) %>%
  DT::datatable()
```

## Cell size

We will now observe the distribution of cell sizes across all images.

```{r cell-size, message=FALSE, fig.height=10, fig.width=10}
dittoPlot(sce, var = "area", group.by = "image",  
          plots = "ridgeplot", color.panel = metadata(sce)$color_vectors$ROIs) 

summary(sce$area)
```

We will now filter cells based on size.

```{r filter-cells}
sum(sce$area <= 7)
sum(sce$area > 400)

sce <- sce[,sce$area > 7]
```

## Expression differences

We will now visualize the expression differences between patients.

```{r expression-differences, fig.height=35}
multi_dittoPlot(sce, vars = rownames(sce)[rowData(sce)$use_channel],
               group.by = "patient_id", plots = c("ridgeplot"), 
               assay = "exprs", 
               color.panel = metadata(sce)$color_vectors$patient_id)
```

We also compute a tSNE and UMAP of all cells to visualize potential sample effects.

```{r dimred}
set.seed(220321)
sce <- runUMAP(sce, subset_row = rowData(sce)$use_channel, exprs_values = "exprs")
```

We can now visualize the patient IDs and expression on the UMAPs.

```{r dimred-viz-patient-indication, message=FALSE, fig.width = 10, fig.height = 10}
dittoDimPlot(sce, var = "patient_id", reduction.use = "UMAP", size = 0.2) + 
    scale_color_manual(values = metadata(sce)$color_vectors$patient_id) +
    ggtitle("Patient ID on UMAP")

dittoDimPlot(sce, var = "indication", reduction.use = "UMAP", size = 0.2) + 
    scale_color_manual(values = metadata(sce)$color_vectors$Indication) +
    ggtitle("Patient ID on UMAP")
```

```{r dimred-viz-expression, fig.height=12, fig.width=20}
p.list <- list()
for(i in rownames(sce)){
  p.list[[i]] <- plotUMAP(sce, colour_by = i, by_exprs_values = "exprs", 
                          point_size = 0.1, point_alpha = 1) 
}

plot_grid(plotlist = p.list, ncol = 7)
```

## Save object

Finally, we will save the modified `SingleCellExperiment` object.

````{r save-RDS}
saveRDS(sce, paste0(mount_path,"IMC/Rout/sce.rds"))
```
