---
title: "Figure 4"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Code to generate the plots of Figure 4.

# set path
the git repo contains a vector which specifies the path to the data folder. please adapt this path in your clone to run the repo automated.
```{r}
mount_path <- readRDS(file = "data/mount_path.rds")
```

## Read data

```{r, message=FALSE}
library(SingleCellExperiment)
library(tidyverse)
library(ggpubr)
library(cowplot)
library(imcRtools)


sce_imc <- readRDS(file.path(mount_path, "IMC/Rout/sce.rds"))
sce_mIF <- readRDS(file.path(mount_path, "mIF/Rout/sce_ROIs.rds"))
```

## Example images

```{r}
getwd()
if (!dir.exists("output/Figure_5/")) dir.create("output/Figure_5/")

cur_IMC <- sce_imc[,sce_imc$sample_id == "IMMUcan_2022_WFLOW_10068868-SPECT-VAR-TIS-01-IMC-01_008"]
cur_mIF <- sce_mIF[,sce_mIF$sample_id == "IMMUcan_2022_WFLOW_10068868-SPECT-VAR-TIS-01-IMC-01_008"]

# Rescale to same x-y dimensions - one pixel in mIF is 0.4962µm
cur_mIF$nucleus.x <- cur_mIF$nucleus.x * 0.4962
cur_mIF$nucleus.y <- cur_mIF$nucleus.y * 0.4962
cur_mIF$TLS.ID <- as.character(cur_mIF$TLS.ID)

# Plot to match Figure 3
(p_4 <- plotSpatial(cur_IMC, img_id = "sample_id", node_color_by = "matched_celltype") + 
  scale_color_manual(values = metadata(sce_imc)$color_vectors$matched_celltype) + 
  xlim(c(600,0)) + ylim(c(0,600)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "none") +
  ggtitle(""))

p_4_with_legend <- plotSpatial(cur_IMC, img_id = "sample_id", node_color_by = "matched_celltype") + 
  scale_color_manual(values = metadata(sce_imc)$color_vectors$matched_celltype) + 
  xlim(c(600,0)) + ylim(c(0,600)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "bottom") +
  ggtitle("")

(p_1 <- plotSpatial(cur_mIF, img_id = "sample_id", node_color_by = "matched_celltype", coords = c("nucleus.x", "nucleus.y")) + 
  scale_color_manual(values = metadata(sce_imc)$color_vectors$matched_celltype) + 
  xlim(c(600,0)) + ylim(c(0,600)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "none") +
  ggtitle(""))

(p_6 <- plotSpatial(cur_IMC, img_id = "sample_id", node_color_by = "CD20_patches") + 
  scale_color_manual(values = c("252" = "red", "257" = "red")) +
  xlim(c(600,0)) + ylim(c(0,600)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "none") +
  ggtitle(""))

(p_3 <- plotSpatial(cur_mIF, img_id = "sample_id", node_color_by = "TLS.ID", coords = c("nucleus.x", "nucleus.y")) + 
  scale_color_manual(values = c("1" = "red")) +
  xlim(c(600,0)) + ylim(c(0,600)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "none") +
  ggtitle(""))

(p_5 <- plotSpatial(cur_IMC, img_id = "sample_id", node_color_by = "tumor_patches") + 
  scale_color_manual(values = c("TRUE" = "sienna", "FALSE" = "grey")) + 
  xlim(c(600,0)) + ylim(c(0,600))+ 
  geom_segment(aes(x = 50, y = 50, xend = 150, yend = 50), lwd = 3, color = "white") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "none") +
  ggtitle(""))

(p_2 <- plotSpatial(cur_mIF, img_id = "sample_id", node_color_by = "tissue.type", coords = c("nucleus.x", "nucleus.y")) + 
  scale_color_manual(values = c("tumor" = "sienna", "stroma" = "grey"))  +
  xlim(c(600,0)) + ylim(c(0,600)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "none") +
  ggtitle(""))

p_final <- plot_grid(p_1, p_3, p_2,p_4,p_6, p_5, nrow = 2, byrow = TRUE) 

ggsave(filename = "output/Figure_5/Figure_5A.pdf", plot = p_final, width = 15, height = 10)
ggsave(filename = "output/Figure_5/Figure_5A_legend.pdf", plot = p_4_with_legend)

```

## Differential abundance testing

```{r, message=FALSE}
library(edgeR)
library(ggrepel)

cur_mIF <- colData(sce_mIF)
cur_IMC <- colData(sce_imc)

# Count cell types for IMC
abundances_IMC <- unclass(table(cur_IMC$matched_celltype, cur_IMC$sample_id))

# Metdata for IMC
meta_IMC <- data.frame(sample_id = colnames(abundances_IMC),
                       patient_id = sub("-", "", str_extract(colnames(abundances_IMC), "[0-9]{8}-")),
                       modality = "IMC")

# Count cell types for mIF
abundances_IF <- unclass(table(cur_mIF$matched_celltype, cur_mIF$sample_id))

# Metadata for mIF
meta_IF <- data.frame(sample_id = colnames(abundances_IF),
                      patient_id = sub("-", "", str_extract(colnames(abundances_IF), "[0-9]{8}-")),
                      modality = "mIF")

colnames(abundances_IMC) <- paste(colnames(abundances_IMC), "IMC")
colnames(abundances_IF) <- paste(colnames(abundances_IF), "mIF")

stopifnot(all.equal(rownames(abundances_IMC), rownames(abundances_IF)))

combined_df <- cbind(abundances_IMC, abundances_IF)
combined_meta <- rbind(meta_IMC, meta_IF)

# DA testing
y.ab <- DGEList(combined_df, samples=combined_meta, group = combined_meta$modality)

# Define the design matrix for the model: Add modality to the model to account for matched design
design <- model.matrix(~sample_id + modality, data = y.ab$samples)

# Estimate NB and QL dispersion
y.ab <- estimateDisp(y.ab, design, trend="none")
summary(y.ab$common.dispersion)

plotBCV(y.ab, cex=1)

fit.ab <- glmQLFit(y.ab, design, robust=TRUE, abundance.trend=FALSE)

summary(fit.ab$var.prior)
summary(fit.ab$df.prior)

plotQLDisp(fit.ab, cex=1)

# Test for differential abundance of cell types 
res <- glmQLFTest(fit.ab, coef=ncol(design))

summary(decideTests(res))
DA <- topTags(res)$table

DA$celltype <- rownames(DA)
DA$sign <- DA$FDR<0.05
DA

# Volcano plot
ggplot(DA) + geom_point(aes(logFC, -log10(FDR), color = sign)) +
  geom_label_repel(aes(logFC, -log10(FDR), label = celltype)) + 
  theme_classic()

# MA plot
(p <- ggplot(DA) + geom_abline(intercept = 0, slope = 0, color = "dark red", lwd = 3) +
  geom_point(aes(logCPM, logFC), size = 7) +
  geom_point(aes(logCPM, logFC, color = sign), size = 5) +
  geom_label_repel(aes(logCPM, logFC, label = celltype), size = 7) + 
  scale_color_manual(values = c("TRUE" = "dark green", "FALSE" = "white"), 
                     name = "Significance") + 
  ylab(c("log2 Fold Change")) + xlab("log2 CPM") +
  ylim(c(-1, 1)) +
  theme_classic(base_size = 25))+
  theme(legend.position = "none")

ggsave(filename = "output/Figure_5/Figure_5B.pdf", plot = p, width = 9, height = 6)

IMC_dat <- as.data.frame(unclass(abundances_IMC))
IMC_dat <- IMC_dat %>%
  mutate(celltype = rownames(IMC_dat)) %>%
  pivot_longer(cols = -celltype) %>%
  mutate(tech = "IMC")
mIF_dat <- as.data.frame(unclass(abundances_IF))
mIF_dat <- mIF_dat %>%
  mutate(celltype = rownames(mIF_dat)) %>%
  pivot_longer(cols = -celltype) %>%
  mutate(tech = "mIF")

dat <- rbind(IMC_dat,mIF_dat)
dat %>%
  ggplot(aes(x=celltype,y = log2(value),fill = as.factor(tech)))+
  geom_boxplot()+
  geom_jitter(position = position_jitterdodge())+
  theme_bw()
dat %>%
  group_by(name) %>%
  mutate(total_cells = sum(value)) %>%
  ggplot(aes(x = tech,y = total_cells))+
  geom_boxplot( )

# some sanity checks
#test <- sce_mIF[,grepl("68868",sce_mIF$sample_id) & sce_mIF$ROI == "005"]
#colnames(test) <- paste0(test$cell.ID)
#dittoSeq::dittoScatterPlot(test,x.var = "CD15",y.var = "CK", color.var = "matched_celltype")
#test <- sce_imc[,grepl("75371",sce_imc$sample_id)]
#dittoSeq::dittoScatterPlot(test,x.var = "CD15",y.var = "CarbonicAnhydrase", color.var = "matched_celltype", assay.x = "exprs", assay.y = "exprs")
#plotSpatial(test, img_id = "sample_id",node_color_by = "celltype")+
#  scale_color_manual(values = metadata(test)$color_vectors$celltype)
```

## Cell phenotype correlation

```{r}
# Match cell types
cur_mIF <- colData(sce_mIF)
cur_IMC <- colData(sce_imc)

cur_mIF$sample_id <- as.factor(cur_mIF$sample_id)
cur_mIF$matched_celltype <- as.factor(cur_mIF$matched_celltype)
cur_IMC$sample_id <- as.factor(cur_IMC$sample_id)
cur_IMC$matched_celltype <- as.factor(cur_IMC$matched_celltype)

frac_IMC <- cur_IMC %>% as_tibble() %>%
  group_by(sample_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

frac_mIF <- cur_mIF %>% as_tibble() %>%
  group_by(sample_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

combined <- left_join(frac_IMC, frac_mIF, by=c("sample_id", "matched_celltype"))
combined$patient_id <- sub("-", "", str_extract(combined$sample_id, "[0-9]{8}-"))

(p_B <- ggplot(combined[combined$matched_celltype == "B",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 0.3)) + xlim(c(0, 0.3)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("B cells") +
  coord_fixed())

(p_BnT <- ggplot(combined[combined$matched_celltype == "BnT",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 0.4)) + xlim(c(0, 0.4)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("BnT cells") +
  coord_fixed())

(p_T <- ggplot(combined[combined$matched_celltype == "T",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 0.6)) + xlim(c(0, 0.6)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("T cells") +
  coord_fixed())

(p_Mac <- ggplot(combined[combined$matched_celltype == "MacCD163",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 0.3)) + xlim(c(0, 0.3)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Macrophages") +
  coord_fixed())

(p_DC <- ggplot(combined[combined$matched_celltype == "DC",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 0.2)) + xlim(c(0, 0.2)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("DC") +
  coord_fixed())

(p_Neutro <- ggplot(combined[combined$matched_celltype == "Neutrophil",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 0.2)) + xlim(c(0, 0.2)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Neutrophils") +
  coord_fixed())

(p_Tumor <- ggplot(combined[combined$matched_celltype == "Tumor",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 0.8)) + xlim(c(0, 0.8)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Tumor") +
  coord_fixed())

(p_other <- ggplot(combined[combined$matched_celltype == "other",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 0.7)) + xlim(c(0, 0.7)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Other cells") +
  coord_fixed())
```
# image based intensity correlation of CD11c

```{r}
exp_IMC <- as.data.frame(t(assay(sce_imc,"exprs")))
exp_mIF <- as.data.frame(t(assay(sce_mIF,"exprs")))
exp_IMC$sample_id <- sce_imc$sample_id
exp_mIF$sample_id <- sce_mIF$sample_id
exp_mIF$matched_celltype <- sce_mIF$matched_celltype

mean_IMC <- exp_IMC %>%
  group_by(sample_id) %>%
  summarise(mean_CD11c_IMC = mean(CD11c))
mean_mif <- exp_mIF %>%
  group_by(sample_id) %>%
  mutate(CD11c = ifelse(is.infinite(CD11c),0,CD11c)) %>%
  summarise(mean_CD11c_mIF = mean(CD11c))

mean_dat <- left_join(mean_IMC,mean_mif,"sample_id")

x <- cor.test(mean_dat$mean_CD11c_IMC,mean_dat$mean_CD11c_mIF,na.action= "na.exclude")
p1<- mean_dat %>%
  ggplot(aes(x = mean_CD11c_IMC,mean_CD11c_mIF))+
  geom_point()+
  theme_bw()+
  ggtitle(paste0("IFQuant mIF pearson's r: ",round(x$estimate, digits = 3)))
p1
# steinbock
mif_steinbock <- readRDS("O:/projects/immucan/processed_data/Panel_1/2022_WORKFLOW/mIF/old/steinbock/Rout/sce.rds")

exp_mIF_st <- as.data.frame(t(assay(mif_steinbock,"exprs")))
exp_mIF_st$sample_id <- mif_steinbock$sample_id
mean_mif <- exp_mIF_st %>%
  group_by(sample_id) %>%
  summarise(mean_CD11c_mIF = mean(CD11c))

mean_dat <- left_join(mean_IMC,mean_mif,"sample_id")

x <- cor.test(mean_dat$mean_CD11c_IMC,mean_dat$mean_CD11c_mIF)
p2 <- mean_dat %>%
  ggplot(aes(x = mean_CD11c_IMC,mean_CD11c_mIF))+
  geom_point()+
  theme_bw()+
  ggtitle(paste0("steinbock mIF pearson's r: ",round(x$estimate, digits = 3)))

cowplot::plot_grid(p1, p2)
test <- t(assay(sce_mIF, "exprs"))

```


## CD20 patch and tumor patch correlation

```{r}
cur_mIF <- colData(sce_mIF)
cur_IMC <- colData(sce_imc)

cur_mIF$sample_id <- as.factor(cur_mIF$sample_id)
cur_mIF$TLS.ID <- as.factor(!is.na(cur_mIF$TLS.ID))
cur_mIF$tissue.type <- as.factor(cur_mIF$tissue.type == "tumor")
cur_IMC$sample_id <- as.factor(cur_IMC$sample_id)
cur_IMC$CD20_patches <- as.factor(!is.na(cur_IMC$CD20_patches))
cur_IMC$tumor_patches <- as.factor(cur_IMC$tumor_patches)

frac_IMC_TLS <- cur_IMC %>% as_tibble() %>%
  group_by(sample_id, CD20_patches, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

frac_mIF_TLS <- cur_mIF %>% as_tibble() %>%
  group_by(sample_id, TLS.ID, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

frac_IMC_tumor <- cur_IMC %>% as_tibble() %>%
  group_by(sample_id, tumor_patches, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

frac_mIF_tumor <- cur_mIF %>% as_tibble() %>%
  group_by(sample_id, tissue.type, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

combined_TLS <- left_join(frac_IMC_TLS, frac_mIF_TLS, by = join_by("sample_id", CD20_patches == TLS.ID))
combined_TLS$patient_id <- sub("-", "", str_extract(combined_TLS$sample_id, "[0-9]{8}-"))

combined_tumor <- left_join(frac_IMC_tumor, frac_mIF_tumor, by = join_by("sample_id", tumor_patches == tissue.type))
combined_tumor$patient_id <- sub("-", "", str_extract(combined_tumor$sample_id, "[0-9]{8}-"))

(p_TLS <- ggplot(combined_TLS[combined_TLS$CD20_patches == TRUE,]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 1)) + xlim(c(0, 1)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("CD20 patches") +
  coord_fixed())

(p_tumor <- ggplot(combined_tumor[combined_tumor$tumor_patches == TRUE,]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_imc)$color_vectors$ROIs, metadata(sce_imc)$color_vectors$patient_id)) +
  ylim(c(0, 1)) + xlim(c(0, 1)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Tumor patches") +
  coord_fixed())

p_final <- plot_grid(p_T, p_B, p_BnT, p_Tumor, p_Mac, p_DC, p_other, p_Neutro, p_TLS, p_tumor, nrow = 2, ncol = 5) 

ggsave(filename = "output/Figure_5/Figure_5C.pdf", plot = p_final, width = 30, height = 15)
```

# Spatial comparison

## Mergin of data

```{r prepare-data}
# Prepare mIF data
cur_mIF <- sce_mIF
colData(cur_mIF) <- colData(cur_mIF)[,c("sample_id", "patient_id", "nucleus.x", "nucleus.y", "matched_celltype")]
cur_mIF$Pos_X <- cur_mIF$nucleus.x
cur_mIF$Pos_Y <- cur_mIF$nucleus.y

cur_mIF$Pos_X <- cur_mIF$Pos_X * 0.4962
cur_mIF$Pos_Y <- cur_mIF$Pos_Y * 0.4962

cur_mIF$nucleus.x <- NULL
cur_mIF$nucleus.y <- NULL

rowData(cur_mIF) <- NULL
assays(cur_mIF) <- list()
metadata(cur_mIF) <- list()

rownames(cur_mIF) <- NULL

colPairs(cur_mIF) <- NULL
cur_mIF$modality <- "mIF"

# Prepare IMC data
cur_IMC <- sce_imc
colData(cur_IMC) <- colData(cur_IMC)[,c("sample_id", "patient_id", "matched_celltype", "Pos_X", "Pos_Y")]

rowData(cur_IMC) <- NULL
assays(cur_IMC) <- list()
metadata(cur_IMC) <- list()

rownames(cur_IMC) <- NULL
reducedDims(cur_IMC) <- NULL
colPairs(cur_IMC) <- NULL
cur_IMC$modality <- "IMC"

# Merge data
all_sce <- cbind(cur_IMC[1:6,], cur_mIF)
all_sce$sample_id_global <- paste0(all_sce$sample_id, "_", all_sce$modality)
```

## Lcross based comparison

```{r laod-libraries, message=FALSE}
library(sf)
library(spatstat)
```

```{r Lcross}
# Example image
cur_sce <- all_sce[,all_sce$sample_id == "IMMUcan_Batch20210921_10082495-SPECT-VAR-TIS-01-IMC-01_004"]
cur_sce <- cur_sce[,cur_sce$matched_celltype == "Tumor"]

plotSpatial(cur_sce, node_color_by = "modality", img_id = "sample_id", flip_y = TRUE)

# Create multipoint pattern
cur_ppp <- ppp(x = cur_sce$Pos_X, y = cur_sce$Pos_Y, 
               xrange = c(0, max(cur_sce$Pos_X)), yrange = c(0, max(cur_sce$Pos_Y)),
               marks = factor(cur_sce$modality, levels = c("IMC", "mIF")))

plot(cur_ppp)

miplot(cur_ppp)
#plot(frypoints(cur_ppp))

cur_L <- Lcross(cur_ppp, "IMC", "mIF")

plot(cur_L)

cur_L_inhom <- Lcross.inhom(cur_ppp, "IMC", "mIF")

plot(cur_L_inhom)

cur_env <- envelope(cur_ppp, Lcross, i = "mIF", j = "IMC", fix.n = TRUE, fix.marks = TRUE)

plot(cur_env)

cur_env <- envelope(cur_ppp, Lcross, i = "IMC", j = "mIF", fix.n = TRUE, fix.marks = TRUE, alternative = "greater")

plot(cur_env)

cur_test <- mad.test(cur_ppp, fun = Lcross, i = "IMC", j = "mIF", fix.n = TRUE, fix.marks = TRUE)
cur_test <- dclf.test(cur_ppp, fun = Lcross, i = "IMC", j = "mIF", fix.n = TRUE, fix.marks = TRUE)

cur_test <- mad.test(cur_ppp, fun = Lcross, i = "IMC", j = "mIF", fix.n = TRUE, fix.marks = TRUE, alternative = "greater")

cur_test <- mad.sigtrace(cur_ppp, fun = Lcross, i = "IMC", j = "mIF", fix.n = TRUE, fix.marks = TRUE)

plot(cur_test)

cur_test <- mad.progress(cur_ppp, fun = Lcross, i = "IMC", j = "mIF", fix.n = TRUE, fix.marks = TRUE)

plot(cur_test)

cur_env_inhom <- envelope(cur_ppp, Lcross.inhom, i = "IMC", j = "mIF")

plot(cur_env_inhom)

cur_test_inhom <- mad.test(cur_ppp, fun = Lcross.inhom, i = "IMC", j = "mIF", rinterval = c(0, 100))
```


```{r automatic}
set.seed(12345)
out <- lapply(unique(all_sce$sample_id), function(x){
  cur_sce <- all_sce[,all_sce$sample_id == x]
  cur_out <- lapply(unique(cur_sce$matched_celltype), function(y){
    cur_sce_2 <- cur_sce[,cur_sce$matched_celltype == y]
    
    if (sum(cur_sce_2$modality == "IMC") < 10 | sum(cur_sce_2$modality == "mIF") < 10) {
      return(NULL)
    }
    
    cur_ppp <- ppp(x = cur_sce_2$Pos_X, y = cur_sce_2$Pos_Y, 
               xrange = c(0, max(cur_sce_2$Pos_X)), yrange = c(0, max(cur_sce_2$Pos_Y)),
               marks = factor(cur_sce_2$modality, levels = c("IMC", "mIF")))
    
    cur_L <- Lcross(cur_ppp, i = "IMC", j = "mIF")
    cur_L_inhom <- Lcross.inhom(cur_ppp, i = "IMC", j = "mIF")
    
    cur_test <- mad.test(cur_ppp, fun = Lcross, i = "IMC", j = "mIF")
    cur_test_inhom <- mad.test(cur_ppp, fun = Lcross.inhom, i = "IMC", j = "mIF")
    
    return(list(test = cur_test, test_inhom = cur_test_inhom, 
                L = cur_L, L_inhom = cur_L_inhom))
  })
  
  names(cur_out) <- unique(cur_sce$matched_celltype)
  
  return(cur_out)
})

names(out) <- unique(all_sce$sample_id)
```

```{r}
cur_summary <- unlist(unlist(out, recursive = FALSE), recursive = FALSE)

cur_p <- lapply(cur_summary[grepl("test$", names(cur_summary))],
                   function(x){
                     return(x$p.value)
                   })

cur_p_inhom <- lapply(cur_summary[grepl("test_inhom$", names(cur_summary))],
                   function(x){
                     return(x$p.value)
                   })

cur_area <- lapply(cur_summary[grepl("L$", names(cur_summary))],
                   function(x){
                     return(sum(x$iso - x$theo))
                   })

cur_area_inhom <- lapply(cur_summary[grepl("L_inhom$", names(cur_summary))],
                   function(x){
                     return(sum(x$iso - x$theo))
                   })

# Check that names match
all.equal(str_split(names(cur_p), "\\.", simplify = TRUE)[,1], str_split(names(cur_p_inhom), "\\.", simplify = TRUE)[,1])
all.equal(str_split(names(cur_p), "\\.", simplify = TRUE)[,1], str_split(names(cur_area), "\\.", simplify = TRUE)[,1])
all.equal(str_split(names(cur_p), "\\.", simplify = TRUE)[,1], str_split(names(cur_area_inhom), "\\.", simplify = TRUE)[,1])

final <- data.frame(p = unlist(cur_p),
                    p_inhom = unlist(cur_p_inhom),
                    area = unlist(cur_area),
                    area_inhom = unlist(cur_area_inhom),
                    sample_id = sub("_", "", str_extract(names(cur_p), "_[0-9]{8}")),
                    celltype = str_split(names(cur_p), "\\.", simplify = TRUE)[,2])

set.seed(123)
p <- ggplot(final) +
  geom_boxplot(aes(celltype, area), outlier.shape = NA) +
  geom_jitter(aes(celltype, area, color = p <= 0.01 & area > 0)) +
  geom_abline(slope = 0, intercept = 0, color = "dark red", linewidth = 1) +
  scale_color_manual(values = c(`TRUE` = "dark blue", `FALSE` = "red"), name = "Significant and larger than 0") +
  theme_classic(base_size = 20) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("") + ylab("Area under L curve")

p

ggsave(filename = "output/Figure_5/Figure_5D.pdf", plot = p, width = 12, height = 7)

final[final$p > 0.01 | final$area < 0,]
```

## Differential connectivity analysis

Run `spicyR`

```{r, fig.height=6, fig.width=8}
library(spicyR)

"sample_id_global" %in% colnames(colData(all_sce))
sum(is.na(colData(all_sce)$sample_id))
sum(is.na(colData(all_sce)$matched_celltype))
sum(is.na(colData(all_sce)$Pos_X))
sum(is.na(colData(all_sce)$Pos_Y))

test <- as.data.frame(colData(all_sce))


spicyTest <- spicy(
  cells=all_sce,
  condition = "modality",
  subject = "sample_id",
  spatialCoordsCols = c("Pos_X", "Pos_Y"),
  window = "square",
  cellTypeCol = "matched_celltype",
  imageIDCol = "sample_id_global",
  Rs = c(20, 50, 100, 150)
)


(p <- signifPlot(
  spicyTest, fdr = T,cutoff = 0.1
))
signifPlot(spicyTest,type = "", comparisonGroup = "IMC")

topPairs(spicyTest, adj = "fdr", cutoff = 0.1)

spicyBoxPlot(results = spicyTest,
             from = "T",
             to = "T")

spicyBoxPlot(results = spicyTest,
             from = "MacCD163",
             to = "MacCD163")

spicyBoxPlot(results = spicyTest,
             from = "Neutrophil",
             to = "other")

spicyBoxPlot(results = spicyTest,
             from = "Neutrophil",
             to = "other")

spicyBoxPlot(results = spicyTest,
             from = "MacCD163",
             to = "Tumor")

spicyBoxPlot(results = spicyTest,
             from = "Tumor",
             to = "MacCD163")

spicyBoxPlot(results = spicyTest,
             from = "T",
             to = "Tumor")

spicyBoxPlot(results = spicyTest,
             from = "Tumor",
             to = "T")

ggsave(filename = "output/Figure_5/Figure_5E_new_2025Feb.pdf", plot = p, width = 7, height = 5)
```

Find example images:

```{r, eval=FALSE}
test1 <- spicyTest$df[spicyTest$df$modality == "IMC",]
test2 <- spicyTest$df[spicyTest$df$modality == "mIF",]

all.equal(sub("_IMC$", "", test1$imageID), sub("_mIF$", "", test2$imageID))

final <- test1[,grepl("__", colnames(test1))] - test2[,grepl("__", colnames(test2))]

test1[order(final[,"T__T"], decreasing = TRUE),c("imageID", "T__T")]

test1[order(final[,"Neutrophil__other"], decreasing = FALSE),c("imageID", "T__T")]

test1[order(final[,"MacCD163__MacCD163"], decreasing = TRUE),c("imageID", "T__T")]

test1[order(final[,"MacCD163__Tumor"], decreasing = FALSE),c("imageID", "T__T")]

test1[order(final[,"T__Tumor"], decreasing = FALSE),c("imageID", "T__T")]
```

# spicyR without TLS images
```{r, fig.height=6, fig.width=8}
library(spicyR)

remove <- unique(all_sce$sample_id)[grepl("_004|_008",unique(all_sce$sample_id))]

spicyTest <- spicy(
  all_sce[,!all_sce$sample_id %in% remove],
  condition = "modality",
  subject = "sample_id",
  spatialCoords = c("Pos_X", "Pos_Y"),
  window = "square",
  cellType = "matched_celltype",
  imageID = "sample_id_global",
  Rs = c(20, 50, 100, 150)
)


(p <- signifPlot(
  spicyTest, fdr = T,cutoff = 0.1
)) 
signifPlot(spicyTest,type = "", comparisonGroup = "IMC")

topPairs(spicyTest, adj = "fdr", cutoff = 0.1)

spicyBoxPlot(results = spicyTest,
             from = "T",
             to = "T")

spicyBoxPlot(results = spicyTest,
             from = "MacCD163",
             to = "MacCD163")

spicyBoxPlot(results = spicyTest,
             from = "Neutrophil",
             to = "other")

spicyBoxPlot(results = spicyTest,
             from = "Neutrophil",
             to = "other")

spicyBoxPlot(results = spicyTest,
             from = "MacCD163",
             to = "Tumor")

spicyBoxPlot(results = spicyTest,
             from = "Tumor",
             to = "MacCD163")

spicyBoxPlot(results = spicyTest,
             from = "T",
             to = "Tumor")

spicyBoxPlot(results = spicyTest,
             from = "Tumor",
             to = "T")
p
#ggsave(filename = "../output/Figure_5/Figure_5E_new.pdf", plot = p, width = 7, height = 5)
```
