---
title: "Figure S2"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: inline
---

Code to generate the plots of Figure S2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# set path
this path should be the directory where all files downloaded from Zenodo are stored in the same structure as on Zenodo
```{r}
mount_path <- readRDS(file = "data/mount_path.rds")
output <- "output/Figure_S2/"
```

## libraries

```{r, message=FALSE}
library(SingleCellExperiment)
library(tidyverse)
library(data.table)
library(cowplot)
library(imcRtools)
library(ggpubr)
library(dittoSeq)
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)
library(scater)
```
# read data

```{r}
cyto_sce <- readRDS(paste0(mount_path,"cytoblocks/Rout/Cleaned_cytoblocks_sce.rds"))

# fix annotation
cyto_sce[,cyto_sce$label == "Apoptotic B-cells"]$label <- "Apoptotic_cells"
names(metadata(cyto_sce)$colour_vectors$cell_types) <- c(names(metadata(cyto_sce)$colour_vectors$cell_types)[c(1:9)],"Apoptotic_cells","CD8 T-cells (non-activated)","NK")
```



# A - activation effect
```{r}
p <- dittoDimPlot(cyto_sce, var = "Activation_status", 
                  reduction.use = "UMAP", size = 0.25,
                  do.label = FALSE)
p
ggsave (filename = "Fig_S2A.png", plot = p, device = "png", path = output, width=200, height=150, units='mm', dpi=600)

```


# B- panel effect
```{r}
p <- dittoDimPlot(cyto_sce, var = "Panel", 
                  reduction.use = "UMAP", size = 0.25,
                  do.label = FALSE)
p
ggsave (filename = "Fig_S2B.png", plot = p, device = "png", path = output, width=200, height=150, units='mm', dpi=600)
```



# C - Marker expression on UMAP

```{r, fig.width=25, fig.height=40}
assay(cyto_sce, "scaled") <- t(scale(t(assay(cyto_sce, "exprs"))))

p.list <- list()
for(i in c("HLADR","GrzB","PDL1","Ecad","LAG3","PD1","Ki67","CD8a","CD4", "CD206","CD20","ICOS")){
  p.list[[i]] <- plotUMAP(cyto_sce, colour_by = i, by_exprs_values = "scaled", 
                          point_size = 0.25, point_alpha = 1) +
    scale_colour_gradient2(name = i, low = "#2166ac", mid = "white", high = "#b2182b",
                           limits = c(-3, 3), oob = squish)+
    theme(legend.title = element_text(size = 20), legend.key.size = unit(8, "mm"))
}

png(file = "output/Figure_S2/Figure_S2C.png",width = 2000 ,height = 1200)
plot_grid(plotlist = p.list, ncol = 4)
dev.off()
```



# D - CV for cell type detection

```{r}
coldata <- as.data.frame(colData(cyto_sce))
# filter out undefined cells for further analysis
act_mean_CV <- coldata %>%
  filter(label != "undefined",
         Activation_status == "activated") %>%
  group_by(Batch, label)%>%
  summarise(n = n()) %>%
  mutate(freq = n/sum(n)) %>%
  group_by(label) %>%
  summarise(CV = sd(freq)/mean(freq)) %>%
  summarise(mean_CV = mean(CV)) %>%
  pull(mean_CV)
p1 <- ggplot(coldata %>%
               filter(label != "undefined",
                      Activation_status == "activated") %>%
             group_by(Batch, label)%>%
               summarise(n = n()) %>%
               mutate(freq = n/sum(n)) %>%
               group_by(label) %>%
               summarise(CV = sd(freq)/mean(freq)),
             # aes(x = label, y = CV across Batches, col = label))+
             aes(x = label, y = CV))+
  geom_point()+
  geom_hline(yintercept = act_mean_CV, linetype = "dashed", color = "red") +
  annotate("text", x = "GrzB", y = act_mean_CV + 0.022, label = paste("Mean:", round(act_mean_CV,2)), hjust = 0.1, color = "red") +
  theme_bw()+
  # scale_color_manual("Labels", values = metadata(sce)$colour_vectors$cell_types) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("activated")

non_act_mean_CV <- coldata %>%
  filter(label != "undefined",
         Activation_status == "non-activated") %>%
  group_by(Batch, label)%>%
  summarise(n = n()) %>%
  mutate(freq = n/sum(n)) %>%
  group_by(label) %>%
  summarise(CV = sd(freq)/mean(freq)) %>%
  summarise(mean_CV = mean(CV)) %>%
  pull(mean_CV)
p2 <- ggplot(coldata %>%
               filter(label != "undefined",
                      Activation_status == "non-activated") %>%
             group_by(Batch, label)%>%
               summarise(n = n()) %>%
               mutate(freq = n/sum(n)) %>%
               group_by(label) %>%
               summarise(CV = sd(freq)/mean(freq)),
             # aes(x = label, y = CV across Batches, col = label))+
             aes(x = label, y = CV))+
  geom_point()+
  geom_hline(yintercept = non_act_mean_CV, linetype = "dashed", color = "red") +
  annotate("text", x = "GrzB", y = non_act_mean_CV + 0.03, label = paste("Mean:", round(non_act_mean_CV,2)), hjust = 0.1, color = "red") +
  theme_bw()+
  # scale_color_manual("Labels", values = metadata(sce)$colour_vectors$cell_types) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("non-activated")


pdf(file = "output/Figure_S2/Fig_S2D.pdf",width = 8, height = 4)
plot_grid(p1,p2,ncol = 2)
dev.off()

```

# E - variation and marker expression


```{r, fig.height=15, fig.width=10}
markers <- list()

# functional markers for T cells
markers[["CD8 T-cells (activated)"]] <- rownames(cyto_sce)[grepl(pattern = "CD27|CD45RA|CD45RO|LAG3|PD1|CD7|GrzB|ICOS|Ki67|TCF7|VISTA",rownames(cyto_sce))]
markers[["CD8 T-cells (non-activated)"]] <- rownames(cyto_sce)[grepl(pattern = "CD27|CD45RA|CD45RO|LAG3|PD1|CD7|GrzB|ICOS|Ki67|TCF7|VISTA",rownames(cyto_sce))]

markers[["T-Helper (activated)"]] <- rownames(cyto_sce)[grepl(pattern = "CD27|CD45RA|CD45RO|LAG3|PD1|CD7|GrzB|FOXP3|ICOS|Ki67|TCF7|VISTA",rownames(cyto_sce))]
markers[["T-Helper (non-activated)"]] <- rownames(cyto_sce)[grepl(pattern = "CD27|CD45RA|CD45RO|LAG3|PD1|CD7|GrzB|FOXP3|ICOS|Ki67|TCF7|VISTA",rownames(cyto_sce))]


# B and BnT cell markers
markers[["B-cells"]] <- rownames(cyto_sce)[grepl(pattern = "CD27|CD45RA|CD45RO|HLADR|CD40|CD20",rownames(cyto_sce))]

# NK markers
markers[["NK"]] <- rownames(cyto_sce)[grepl(pattern = "CD7|GrzB|CD16$|CD45RA|CD45RO",rownames(cyto_sce))]

# tumor markers
markers[["PDL-1 positive Epithelial cells"]] <- rownames(cyto_sce)[grepl(pattern = "Ecad|HLADR|CarbonicAnhydrase|cleavedPARP|PDL1|IDO1|CD15|Ki67",rownames(cyto_sce))]

markers[["Ecad positive Epithelial cells"]] <- rownames(cyto_sce)[grepl(pattern = "Ecad|HLADR|CarbonicAnhydrase|cleavedPARP|PDL1|IDO1|CD15|Ki67",rownames(cyto_sce))]

# myeloid markers
markers[["Macrophages"]] <- rownames(cyto_sce)[grepl(pattern = "CD68|CD11c|PDL1|ICOS|VISTA|CD40|CD14|CD206|IDO1|CD33|CD16|CD163",rownames(cyto_sce))]
markers[["Myeloid"]] <- rownames(cyto_sce)[grepl(pattern = "CD68|CD11c|PDL1|ICOS|VISTA|CD40|CD14|CD206|IDO1|CD33|CD16|CD163",rownames(cyto_sce))]

# Apoptotic
markers[["Apoptotic B-cells"]] <- rownames(cyto_sce)[grepl(pattern = "cleavedPARP",rownames(cyto_sce))]


# separate by activation status
cyto_sce_act <- cyto_sce[,cyto_sce$Activation_status == "activated"]
cyto_sce_non_act <- cyto_sce[,cyto_sce$Activation_status == "non-activated"]




```

## activated cells 

```{r}
# compute mean expression of marker per batch and cluster
# Activated
library(reshape2)

# extract expression matrix as dataframe and add label and batch information
expr_df_act <- as.data.frame(t(assay(cyto_sce_act, "exprs")))
expr_df_act$label <- cyto_sce_act$label
expr_df_act$Batch <- cyto_sce_act$Batch

# reshape dataframe so it fits our needs
expr_df_act <- expr_df_act %>%
  group_by(Batch,label)%>%
  summarise(across(everything(), mean),
            .groups = 'drop')

expr_df_act <- melt(expr_df_act, id.vars=c('Batch',"label"), 
                  measure.vars=rownames(cyto_sce),
                  na.rm = F)

# select for relevant markers per cell type
for (lbl in unique(expr_df_act$label)){
  expr_df_act <- expr_df_act[!((expr_df_act$label == lbl) & !(expr_df_act$variable %in% markers[[lbl]])),]
}

```

## non-activated cells

```{r}
# extract expression matrix as dataframe and add label and batch information
expr_df_non_act <- as.data.frame(t(assay(cyto_sce_non_act, "exprs")))
expr_df_non_act$label <- cyto_sce_non_act$label
expr_df_non_act$Batch <- cyto_sce_non_act$Batch

# reshape dataframe so it fits our needs
expr_df_non_act <- expr_df_non_act %>%
  group_by(Batch,label)%>%
  summarise(across(everything(), mean),
            .groups = 'drop')

expr_df_non_act <- melt(expr_df_non_act, id.vars=c('Batch',"label"), 
                  measure.vars=rownames(cyto_sce),
                  na.rm = F)

# select for relevant markers per cell type
for (lbl in unique(expr_df_non_act$label)){
  expr_df_non_act <- expr_df_non_act[!((expr_df_non_act$label == lbl) & !(expr_df_non_act$variable %in% markers[[lbl]])),]
}

```

##  plot

```{r}
cv_act <- expr_df_act %>%
  group_by(label,variable) %>%
  summarise(CV = sd(value)/mean(value))

# fit linear model
res <- expr_df_act %>%
  group_by(label,variable) %>%
  mutate(CV = sd(value)/mean(value)) %>%
  group_by(variable) %>%
  summarise(mean_value = mean(value),
            mean_CV = mean(CV))

fit <- lm(formula = mean_CV ~ mean_value, data = res)

r_square <- summary(fit)$r.squared


# plot data

p <- expr_df_act %>%
  group_by(label,variable) %>%
  mutate(CV = sd(value)/mean(value)) %>%
  group_by(variable) %>%
  summarise(mean_value = mean(value),
            mean_CV = mean(CV)) %>%
  ggplot(aes(x = mean_value, y = mean_CV))+
  geom_smooth(method = lm, formula = y ~ x ,)+
  geom_point(aes(col =variable))+
  stat_regline_equation(aes(label =  ..rr.label..))+
  theme_bw()

p
ggsave (filename = "Fig_S2E.pdf", plot = p, device = "pdf", path = output, width = 16, height = 10, units = "cm")
```

# G CV - non_activated

```{r}
cv_non_act <- expr_df_non_act %>%
  group_by(label,variable) %>%
  summarise(CV = sd(value)/mean(value))

marker_cv_non_act <- cv_non_act%>%
  group_by(variable) %>%
  summarise(mean = mean(CV), sd = sd(CV), n = n()) %>%
  mutate(upper = mean + sd, lower = mean - sd)


ymax <- max(marker_cv_non_act$mean)
ymin <- min(marker_cv_non_act$mean)
ymean <- mean(marker_cv_non_act$mean)
p <- ggplot(marker_cv_non_act, aes(x = variable))+
  # geom_errorbar(aes(ymin=lower, ymax =upper))+
  geom_point(aes(y = mean)) +
  # geom_text(aes(label=n, y=upper+0.02)) +
  geom_hline(yintercept = ymax, linetype = "dashed", color = "red") +
  annotate("text", x = "cleavedPARP", y = ymax - 0.022, label = paste("Max:", round(ymax,2)), hjust = 1.1, color = "red") +
  geom_hline(yintercept = ymin, linetype = "dashed", color = "red") +
  annotate("text", x = "CD16", y = ymin + 0.022, label = paste("Min:", round(ymin,2)), hjust = -0.1, color = "red") +
  geom_hline(yintercept = ymean, linetype = "dashed", color = "red") +
  annotate("text", x = "GrzB", y = ymean + 0.022, label = paste("Mean:", round(ymean,2)), hjust = 0.1, color = "red") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Marker")+
  ylab("mean CV")
p
#ggsave (filename = "Fig_S2G_non_activated.pdf", plot = p, device = "pdf", path = output, width = 12, height = 8, units = "cm")
```

