---
title: "Figure S4"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: inline
---

Code to generate the plots of Figure S4.
for a more detailed description and training of the random forest the classifier for IMMUcan panel 1 please also see this git repo [here](https://github.com/ImmucanWP7/IMMUcan_IMC_panel_1_cell_type_classifier)

In this script we only load the pretrained classifier to generated Figure S4

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Code to generate the plots of Figure 4.

# set path
the git repo contains a vector which specifies the path to the data folder. please adapt this path in your clone to run the repo automated.
```{r}
mount_path <- readRDS(file = "data/mount_path.rds")
output <- "output/Figure_S4"
```


load libraries
```{r}
library(SingleCellExperiment)
library(caret)
library(doParallel)
library(tidyverse)
library(ggplot2)
```

# load data
training data consists of labels from the first round of labelling (V1) and a second round of labelling (V3)

```{r}
sce <- readRDS(paste0(mount_path,"IMC/Rout/sce_labelled_V1.rds"))
sce_new <- readRDS(paste0(mount_path,"IMC/Rout/sce_labelled_V3.rds"))
```

## Exclude DCs with low CD11c expression

We observed an over-estimation of DCs in the trials and will now exclude DCs with low CD11c. 

```{r exclude-DCs}
ggplot(data.frame(CD11c = assay(sce[,sce$cell_labels == "DC"], "exprs")["CD11c",],
                  CD68 = assay(sce[,sce$cell_labels == "DC"], "exprs")["CD68",])) +
  geom_point(aes(CD11c, CD68))

sce$cell_labels[sce$cell_labels == "DC" & (assay(sce, "exprs")["CD11c",] < 2 | assay(sce, "exprs")["CD68",] > 3)] <- "unlabelled"

ggplot(data.frame(CD11c = assay(sce[,sce$cell_labels == "DC"], "exprs")["CD11c",],
                  CD68 = assay(sce[,sce$cell_labels == "DC"], "exprs")["CD68",])) +
  geom_point(aes(CD11c, CD68))

ggplot(data.frame(CD11c = assay(sce_new[,sce_new$cell_labels == "DC"], "exprs")["CD11c",],
                  CD68 = assay(sce_new[,sce_new$cell_labels == "DC"], "exprs")["CD68",])) +
  geom_point(aes(CD11c, CD68))

sce_new$cell_labels[sce_new$cell_labels == "DC" & (assay(sce_new, "exprs")["CD11c",] < 2 | assay(sce_new, "exprs")["CD68",] > 3)] <- "unlabelled"

ggplot(data.frame(CD11c = assay(sce_new[,sce_new$cell_labels == "DC"], "exprs")["CD11c",],
                  CD68 = assay(sce_new[,sce_new$cell_labels == "DC"], "exprs")["CD68",])) +
  geom_point(aes(CD11c, CD68))

lab_sce <- sce[,sce$cell_labels != "unlabelled"]
unlab_sce <- sce[,sce$cell_labels == "unlabelled"]

lab_sce_new <- sce_new[,sce_new$cell_labels != "unlabelled"]
unlab_sce_new <- sce_new[,sce_new$cell_labels == "unlabelled"]

table(lab_sce$cell_labels)
table(lab_sce_new$cell_labels)
```

Next, we will first split the labelled data into training and test (validation) data at a ratio of 80/20 using images as grouping level.

```{r split-data-images-DCs}
set.seed(221107)
trainIndex <- groupKFold(group = factor(lab_sce$sample_id), k = 5)
train_sce <- lab_sce[,trainIndex$Fold1]
test_sce <- lab_sce[,-trainIndex$Fold1]

table(train_sce$cell_labels)
table(test_sce$cell_labels)

set.seed(230410)
trainIndex_new <- groupKFold(group = factor(lab_sce_new$sample_id), k = 5)
train_sce_new <- lab_sce_new[,trainIndex_new$Fold1]
test_sce_new <- lab_sce_new[,-trainIndex_new$Fold1]

table(train_sce_new$cell_labels)
table(test_sce_new$cell_labels)

cur_names <- Reduce(intersect, x = list(colnames(colData(train_sce)), colnames(colData(train_sce_new))))
colData(train_sce) <- colData(train_sce)[,cur_names]
colData(test_sce) <- colData(test_sce)[,cur_names]
colData(train_sce_new) <- colData(train_sce_new)[,cur_names]
colData(test_sce_new) <- colData(test_sce_new)[,cur_names]

reducedDims(train_sce) <- NULL
reducedDims(test_sce) <- NULL
reducedDims(train_sce_new) <- NULL
reducedDims(test_sce_new) <- NULL

train_sce$classifier <- "v1"
test_sce$classifier <- "v1"
train_sce_new$classifier <- "v3"
test_sce_new$classifier <- "v3"

final_train <- cbind(train_sce, train_sce_new)
final_test <- cbind(test_sce, test_sce_new)

table(final_train$cell_labels, final_train$classifier)
table(final_test$cell_labels, final_test$classifier)
```

Note: we will load the pretrained classifier here

```{r accuracy-images-DCs}
rffit <- readRDS(paste0(mount_path,"IMC/Rout/rf_images_DCfix.rds"))

ggplot(rffit) + 
  geom_errorbar(data = rffit$results,
                aes(ymin = Accuracy - AccuracySD,
                    ymax = Accuracy + AccuracySD),
                width = 0.4)
```

We can also compute the confusion matrix:

```{r confusion-matrix-images-DCs}
confusionMatrix(rffit)
```

We will also look at the variable importance.

```{r variable-importance-images-DCs, fig.height = 15}
cur_varImp <- varImp(rffit)
plot(cur_varImp)
```

## Old test data

Finally, we will validate the model using the test data.

```{r model-testing-images-DCs-old-data}
# Add dummy variables
cur_sce <- final_test[,final_test$classifier == "v1"]
cur_mat <- t(assay(cur_sce, "exprs")[!grepl("DNA|Histone", rownames(cur_sce)),])

dummies <- dummyVars(SampleId ~ Indication, data = colData(cur_sce))
all_dummies <- predict(dummies, newdata = colData(cur_sce))

cur_mat <- cbind(cur_mat, all_dummies)

cur_pred <- predict(rffit, 
                    newdata = cur_mat)

cm <- confusionMatrix(data = cur_pred, reference = factor(cur_sce$cell_labels), mode = "everything")
cm

data.frame(cm$byClass) %>%
  mutate(class = sub("Class: ", "", rownames(cm$byClass))) %>%
  ggplot() + 
  geom_point(aes(1 - Specificity, Sensitivity, 
                 size = Detection.Rate,
                 fill = class),
             shape = 21) + 
  scale_fill_manual(values = metadata(sce)$colour_vectors$cell_types) + 
  theme_bw() + 
  ylab("Sensitivity (TPR)") +
  xlab("1 - Specificity (FPR)")
```

We will also observe the distribution of classification probabilities per image and class:

```{r prediciton-probability-images-DCs-old-data, fig.width = 15}
cur_pred <- predict(rffit, 
                    newdata = cur_mat, 
                    type = "prob")
cur_pred$truth <- factor(cur_sce$cell_labels)

cur_pred %>%
  pivot_longer(cols = B:Tumor) %>%
  ggplot() +
  geom_boxplot(aes(x = name, y = value, fill = name), outlier.size = 0.5) +
  facet_wrap(. ~ truth) + 
  scale_fill_manual(values = metadata(sce)$colour_vectors$cell_types) +
  theme(panel.background = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 1))
```

## New test data

Now we specifically test the classifier on the newly labelled cells.

```{r model-testing-images-DCs-new-data}
# Add dummy variables
cur_sce <- final_test[,final_test$classifier == "v3"]
cur_mat <- t(assay(cur_sce, "exprs")[!grepl("DNA|Histone", rownames(cur_sce)),])

dummies <- dummyVars(SampleId ~ Indication, data = colData(cur_sce))
all_dummies <- predict(dummies, newdata = colData(cur_sce))
all_dummies <- cbind(all_dummies, matrix(0, nrow = nrow(all_dummies), dimnames = list(rownames(all_dummies), "IndicationGI")))

cur_mat <- cbind(cur_mat, all_dummies)

cur_pred <- predict(rffit, 
                    newdata = cur_mat)
cur_ref <- factor(cur_sce$cell_labels, levels = levels(cur_pred))

cm <- confusionMatrix(data = cur_pred, reference = cur_ref, mode = "everything")
cm

data.frame(cm$byClass) %>%
  mutate(class = sub("Class: ", "", rownames(cm$byClass))) %>%
  ggplot() + 
  geom_point(aes(1 - Specificity, Sensitivity, 
                 size = Detection.Rate,
                 fill = class),
             shape = 21) + 
  scale_fill_manual(values = metadata(sce)$colour_vectors$cell_types) + 
  theme_bw() + 
  ylab("Sensitivity (TPR)") +
  xlab("1 - Specificity (FPR)")
```

We will also observe the distribution of classification probabilities per image and class:

```{r prediciton-probability-images-DCs-new-data, fig.width = 7}
cur_pred <- predict(rffit, 
                    newdata = cur_mat, 
                    type = "prob")
cur_pred$truth <- factor(cur_sce$cell_labels)

cur_pred %>%
  pivot_longer(cols = B:Tumor) %>%
  ggplot() +
  geom_boxplot(aes(x = name, y = value, fill = name), outlier.size = 0.5) +
  facet_wrap(. ~ truth) + 
  scale_fill_manual(values = metadata(sce)$colour_vectors$cell_types) +
  theme(panel.background = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 1))
```

## combined test data

We will check the model performance on the complete test data

```{r model-testing-images-DCs-new-data}
# Add dummy variables
cur_sce <- final_test
cur_mat <- t(assay(cur_sce, "exprs")[!grepl("DNA|Histone", rownames(cur_sce)),])

dummies <- dummyVars(SampleId ~ Indication, data = colData(cur_sce))
all_dummies <- predict(dummies, newdata = colData(cur_sce))
all_dummies <- cbind(all_dummies, matrix(0, nrow = nrow(all_dummies), dimnames = list(rownames(all_dummies), "IndicationGI")))

cur_mat <- cbind(cur_mat, all_dummies)

cur_pred <- predict(rffit, 
                    newdata = cur_mat)
cur_ref <- factor(cur_sce$cell_labels, levels = levels(cur_pred))

cm <- confusionMatrix(data = cur_pred, reference = cur_ref, mode = "everything")
cm

p <- data.frame(cm$byClass) %>%
  mutate(class = sub("Class: ", "", rownames(cm$byClass))) %>%
  ggplot() + 
  geom_point(aes(1 - Specificity, Sensitivity, 
                 size = Detection.Rate,
                 fill = class),
             shape = 21) + 
  scale_fill_manual(values = metadata(sce)$colour_vectors$cell_types) + 
  theme_bw() + 
  scale_size(range = c(1,12))+
  ylab("Sensitivity (TPR)") +
  xlab("1 - Specificity (FPR)")
p

ggsave (filename = "Fig_S4A.pdf", plot = p, device = "pdf", path = output, width = 8, height = 6)
```

```{r prediciton-probability-images-DCs-new-data, fig.width = 15}
cur_pred <- predict(rffit, 
                    newdata = cur_mat, 
                    type = "prob")
cur_pred$truth <- factor(cur_sce$cell_labels)

p <- cur_pred %>%
  pivot_longer(cols = B:Tumor) %>%
  ggplot() +
  geom_boxplot(aes(x = name, y = value, fill = name), outlier.size = 0.5) +
  facet_wrap(. ~ truth) + 
  scale_fill_manual(values = metadata(sce)$colour_vectors$cell_types) +
  theme(panel.background = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 1))+
  ylab("probability")+
  xlab("cell type")
p
ggsave (filename = "Fig_S4B.png", plot = p, device = "png", path = output, width = 35, height = 22,units = "cm",dpi = 1200)
```

