---
title: "Figure S5"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: inline
---

Code to generate the plots of Figure S5.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Code to generate the plots of Figure 4.

# set path
the git repo contains a vector which specifies the path to the data folder. please adapt this path in your clone to run the repo automated.
```{r}
mount_path <- readRDS(file = "data/mount_path.rds")
output <- "output/Figure_S5"
```

## Read data

```{r, message=FALSE}
library(SingleCellExperiment)
library(tidyverse)
library(ggpubr)
library(cowplot)
library(imcRtools)


sce_IMC <- readRDS(file.path(mount_path, "IMC/Rout/sce.rds"))
sce_mIF <- readRDS(file.path(mount_path, "mIF/Rout/sce_ROIs.rds"))
```

# A - Number of cells

```{r}
if (!dir.exists("output/Figure_S5/")) dir.create("output/Figure_S5/")

no_IMC <- colData(sce_IMC) %>% as_tibble() %>%
  group_by(sample_id, patient_id) %>%
  summarize(count = n())
no_mIF <- colData(sce_mIF) %>% as_tibble() %>%
  group_by(sample_id, patient_id) %>%
  summarize(count = n())

combined_df <- left_join(no_IMC, no_mIF, by = c("sample_id", "patient_id"))
combined_df$patient_id <- sub("-", "", str_extract(combined_df$sample_id, "[0-9]{8}-"))

p <- ggplot(combined_df) + 
  geom_point(aes(count.x, count.y, color = patient_id), size = 2) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red") +
  geom_smooth(aes(count.x, count.y), method='lm', formula= y~x, fullrange=TRUE) + 
  theme_minimal(base_size = 20) + xlab("# Cells in IMC") + ylab("# Cells in mIF") +
  coord_fixed() + 
  ylim(c(0, 5000)) + xlim(c(0, 5000)) +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$patient_id, name = "Patient ID")

p
ggsave(filename = "FigS5_A.pdf",path = output, plot = p, width = 10, height = 7)
```

# B - cellular area

```{r}
mif_steinbock <- readRDS("O:/projects/immucan/processed_data/Panel_1/2022_WORKFLOW/mIF/old/steinbock/Rout/sce.rds")
IMC_dat <- data.frame(IMC_cell_size = sce_IMC$area)
mIF_dat <- data.frame(IF_cell_size = sce_mIF$cell.area)
mIF_steinbock_dat <- data.frame(IF_cell_size = mif_steinbock$area)

# correct for difference in resolution: one pixel in mIF is 0.4962µm
mIF_steinbock_dat$IF_cell_size <- mIF_steinbock_dat$IF_cell_size*0.4962^2

mean_IMC <- mean(IMC_dat$IMC_cell_size)
mean_IF <- mean(mIF_dat$IF_cell_size )
mean_IF_steinbock <- mean(mIF_steinbock_dat$IF_cell_size)


cur_dat <- data.frame(IMC_cell_size = sce_IMC$area)

mean_IMC <- mean(cur_dat$IMC_cell_size)
mean_IF <- mean(sce_mIF$cell.area)

p <- cur_dat %>%
  ggplot(aes(y= IMC_cell_size))+
  geom_density(col= "blue", size = 2)+
  coord_flip()+
  theme_bw()+
  geom_density(data = mIF_dat ,aes(y = IF_cell_size),col= "red", size = 2)+
  geom_density(data = mIF_steinbock_dat ,aes(y = IF_cell_size),col= "green", size = 2)+
  annotate("text", x = 0.02, y = 392 , label = bquote("IMC: average" ~ .(round(mean_IMC)) ~  µm^2), color = "blue", size = 5)+
  annotate("text", x = 0.022, y = 400 , label =  bquote("mIF: average" ~ .(round(mean_IF)) ~  µm^2), color = "red", size = 5)+
  annotate("text", x = 0.024, y = 400 , label =bquote("mIF steinbock: average" ~ .(round(mean_IF_steinbock)) ~  µm^2), color = "green", size = 5)+
  ylab("cell size")+
  xlab("density estimate")

p

ggsave(filename = "FigS5_B.pdf",path= output, plot = p, width = 7, height = 5)
area = pi*r^2
area <- 78
r <- sqrt(area/pi)
r
cur_sce <- sce_mIF[,sce_mIF$sample_id == "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_001"]
cur_sce_2 <- sce_IMC[,sce_IMC$sample_id == "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_001"]

sum(cur_sce$cell.area)
sum(cur_sce_2$area)
plotSpatial(cur_sce,img_id = "sample_id",coords = c("nucleus.x","nucleus.y"),node_color_by = "matched_celltype")
plotSpatial(cur_sce_2,img_id = "sample_id",node_color_by = "matched_celltype")
```




# Supporting plots for DA analysis

```{r, fig.height=10, fig.width=10}
cur_mIF <- colData(sce_mIF)
cur_IMC <- colData(sce_IMC)

# Proportions of cell types for IMC
cur_IMC$matched_celltype <- factor(cur_IMC$matched_celltype)
prop_IMC <- cur_IMC %>% 
  as.data.frame() %>%
  group_by(sample_id, patient_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  group_by(sample_id) %>%
  mutate(prop = count/sum(count))

# Proportions of cell types for mIF
cur_mIF$matched_celltype <- factor(cur_mIF$matched_celltype)
prop_mIF <- cur_mIF %>% 
  as.data.frame() %>%
  group_by(sample_id, patient_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  group_by(sample_id) %>%
  mutate(prop = count/sum(count))

prop_IMC$modality <- "IMC"
prop_mIF$modality <- "mIF"

out <- rbind(prop_IMC, prop_mIF)

(p <- ggplot(out) + 
  geom_boxplot(aes(modality, prop), outlier.shape = NA) +
  geom_point(aes(modality, prop, color = patient_id)) +
  geom_line(aes(modality, prop, group = sample_id)) +
  facet_wrap(~matched_celltype, nrow = 1, scales = "free")  +
  theme_classic(base_size = 20) +
  ylab("Proportion of cell phenotypes") +
  xlab("Modality") +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$patient_id, "Patient ID"))
p
ggsave(filename = "FigS5_C.pdf",path=output, plot = p, height = 5, width = 17)
```

Example images

```{r, eval=FALSE}
library(cytomapper)

(p1 <- plotSpatial(sce_IMC[,sce_IMC$sample_id == "IMMUcan_2022_WFLOW_10073140-SPECT-VAR-TIS-01-IMC-01_004"], 
            img_id = "sample_id", node_color_by = "matched_celltype") +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype) +
    theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "none") + ggtitle(""))

(p2 <- plotSpatial(sce_mIF[,sce_mIF$sample_id == "IMMUcan_2022_WFLOW_10073140-SPECT-VAR-TIS-01-IMC-01_004"], 
            img_id = "sample_id", node_color_by = "matched_celltype", coords = c("nucleus.x", "nucleus.y")) +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype) +
    theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA)) + ggtitle(""))

#ggsave(filename = "../output/Figure_S5/FigS5_D1.pdf", plot = p1, height = 10, width = 10)
#ggsave(filename = "../output/Figure_S5/FigS5_D2.pdf", plot = p2, height = 10, width = 10)
```


# whole slide comparison
we will correlate the fractions of cell types from mIF and IMC. we will therefore exclude ROI4 (TLS) which will otherwise falsify the results because they add many B cells to the IMC for a specific patient
```{r, fig.height=10, fig.width=10}
mIF_sce <- readRDS(paste0(mount_path,"mIF/Rout/sce_whole_slide.rds"))

mIF_sce
unique(mIF_sce$patient_id)
unique(sce_IMC$patient_id)

cur_imc <- sce_IMC[,sce_IMC$ROI != 4]

mIF <- as.data.frame(colData(mIF_sce))
IMC <- as.data.frame(colData(sce_IMC))
mIF <- mIF %>%
  group_by(patient_id) %>%
  count(matched_celltype, name = "ct_count_mIF") %>%
  mutate(total_cell_count_mIF = sum(ct_count_mIF),
         frac_cellt_mIF = ct_count_mIF/total_cell_count_mIF)

IMC <- IMC %>%
  filter(! ROI %in% c(4,8)) %>%
  group_by(patient_id) %>%
  count(matched_celltype, name = "ct_count_IMC") %>%
  mutate(total_cell_count_IMC = sum(ct_count_IMC),
         frac_cellt_IMC = ct_count_IMC/total_cell_count_IMC)

dat <- left_join(IMC, mIF, c("patient_id","matched_celltype"))

dat <- dat %>%
  pivot_wider(id_cols = patient_id,names_from = matched_celltype,values_from = c(frac_cellt_IMC,frac_cellt_mIF))
cor <- psych::corr.test(dat[,grepl("_IMC",colnames(dat))],dat[,grepl("mIF",colnames(dat))],  method = "pearson")

cur_dat <- as.data.frame(cor$r)
cur_dat$variable <- rownames(cur_dat)
dat_long <- reshape2::melt(cur_dat,id.vars="variable")
colnames(dat_long) <- c("cell_types_IMC","cell_types_mIF","correlation")

#get p-value matrix
p_dat <- as_tibble(cor$p)
p_dat$variable <- rownames(cur_dat)
pdat_long <- reshape2::melt(p_dat,id.vars="variable")
colnames(pdat_long) <- c("cell_types_IMC","cell_types_mIF","p_adj")

dat_long$p_adj <- pdat_long$p_adj

dat_long <- dat_long %>%
  mutate(sig = ifelse(p_adj <= 0.001 & p_adj > 0.0001,0.001,p_adj))

dat_long <- dat_long %>%
  mutate(sig = case_when(p_adj <= 0.0001 ~ "< 0.0001",
                         p_adj <= 0.001 & p_adj > 0.0001 ~ "< 0.001",
                         p_adj <= 0.01 & p_adj > 0.001 ~ "< 0.01",
                         p_adj <= 0.1 & p_adj > 0.01 ~ "< 0.1",
                         p_adj >0.1 ~ "ns"))

cor_filter <- 0.0
library(ggnewscale)
p <- dat_long %>%
   filter(abs(correlation) > cor_filter) %>%
   ggplot(aes(x=factor(cell_types_IMC, levels=sort(unique(dat_long$cell_types_IMC))),y = factor(cell_types_mIF, levels=sort(levels(dat_long$cell_types_mIF))),col = correlation, size= -log10(p_adj)))+
    #geom_tile(aes(fill = sig))+
    #theme_bw()+
    geom_point()+
    theme_bw()+
    scale_color_gradient2(low =  "blue",mid = "white",high = "red", name = expression("Pearson's" ~ rho))+
    #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    #      panel.background = element_blank(), axis.line = element_line(colour = "black"))+
    #theme_bw()+
    #scale_color_gradient(low = "white",   high = "red")+
    scale_size(range = c(1,8))+
    new_scale_colour() +
    geom_point(data = dat_long %>% filter(sig <0.01, abs(correlation) > cor_filter),shape = 1,aes(size = -log10(p_adj), col = sig), stroke =2.5)+
    scale_color_manual(values = c("TRUE" = "black","FALSE" = "white"), name = "FDR < 0.1")+
    theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1,size = 20))+
    theme(axis.text.y = element_text(size = 20))+
    theme(legend.text = element_text(size = 20))+
    theme(legend.title = element_text(size = 20))
p

ggsave(filename = paste0(output,"/wholeslide_mIF_IMC_correlation.pdf"),plot = p,width = 12, height = 8)

dat %>%
  ggplot(aes(x = frac_cellt_IMC_MacCD163,y = frac_cellt_mIF_MacCD163))+
  geom_point()+
  geom_abline(slope = 1,intercept = 0)+
  coord_cartesian(xlim = c(0,1),ylim = c(0,1))+
  theme_bw()
cor(dat$frac_cellt_mIF_MacCD163,dat$frac_cellt_IMC_MacCD163)


```

```{r, fig.height=10, fig.width=10}
mIF_sce <- readRDS(paste0(mount_path,"mIF/Rout/sce_whole_slide.rds"))

mIF_sce
unique(mIF_sce$patient_id)
unique(sce_IMC$patient_id)

cur_imc <- sce_IMC[,sce_IMC$ROI != 4]

mIF <- as.data.frame(colData(sce_mIF))
IMC <- as.data.frame(colData(sce_IMC))
mIF <- mIF %>%
  group_by(patient_id) %>%
  count(matched_celltype, name = "ct_count_mIF") %>%
  mutate(total_cell_count_mIF = sum(ct_count_mIF),
         frac_cellt_mIF = ct_count_mIF/total_cell_count_mIF)

IMC <- IMC %>%
  filter(! ROI %in% c(4,8)) %>%
  group_by(patient_id) %>%
  count(matched_celltype, name = "ct_count_IMC") %>%
  mutate(total_cell_count_IMC = sum(ct_count_IMC),
         frac_cellt_IMC = ct_count_IMC/total_cell_count_IMC)

dat <- left_join(IMC, mIF, c("patient_id","matched_celltype"))

dat <- dat %>%
  pivot_wider(id_cols = patient_id,names_from = matched_celltype,values_from = c(frac_cellt_IMC,frac_cellt_mIF))
cor <- psych::corr.test(dat[,grepl("_IMC",colnames(dat))],dat[,grepl("mIF",colnames(dat))],  method = "pearson")

cur_dat <- as.data.frame(cor$r)
cur_dat$variable <- rownames(cur_dat)
dat_long <- reshape2::melt(cur_dat,id.vars="variable")
colnames(dat_long) <- c("cell_types_IMC","cell_types_mIF","correlation")

#get p-value matrix
p_dat <- as_tibble(cor$p)
p_dat$variable <- rownames(cur_dat)
pdat_long <- reshape2::melt(p_dat,id.vars="variable")
colnames(pdat_long) <- c("cell_types_IMC","cell_types_mIF","p_adj")

dat_long$p_adj <- pdat_long$p_adj

dat_long <- dat_long %>%
  mutate(sig = ifelse(p_adj <= 0.001 & p_adj > 0.0001,0.001,p_adj))

dat_long <- dat_long %>%
  mutate(sig = case_when(p_adj <= 0.0001 ~ "< 0.0001",
                         p_adj <= 0.001 & p_adj > 0.0001 ~ "< 0.001",
                         p_adj <= 0.01 & p_adj > 0.001 ~ "< 0.01",
                         p_adj <= 0.1 & p_adj > 0.01 ~ "< 0.1",
                         p_adj >0.1 ~ "ns"))

cor_filter <- 0.0
p <- dat_long %>%
   filter(abs(correlation) > cor_filter) %>%
   ggplot(aes(x=factor(cell_types_IMC, levels=sort(unique(dat_long$cell_types_IMC))),y = factor(cell_types_mIF, levels=sort(levels(dat_long$cell_types_mIF))),col = correlation, size= -log10(p_adj)))+
    #geom_tile(aes(fill = sig))+
    #theme_bw()+
    geom_point()+
    theme_bw()+
    scale_color_gradient2(low =  "blue",mid = "white",high = "red", name = expression("Pearson's" ~ rho))+
    #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    #      panel.background = element_blank(), axis.line = element_line(colour = "black"))+
    #theme_bw()+
    #scale_color_gradient(low = "white",   high = "red")+
    scale_size(range = c(1,8))+
    new_scale_colour() +
    geom_point(data = dat_long %>% filter(sig <0.01, abs(correlation) > cor_filter),shape = 1,aes(size = -log10(p_adj), col = sig), stroke =2.5)+
    scale_color_manual(values = c("TRUE" = "black","FALSE" = "white"), name = "FDR < 0.1")+
    theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1,size = 20))+
    theme(axis.text.y = element_text(size = 20))+
    theme(legend.text = element_text(size = 20))+
    theme(legend.title = element_text(size = 20))
p

ggsave(filename = paste0(output,"/matched_mIF_IMC_correlation.pdf"),plot = p,width = 12, height = 8)

dat %>%
  ggplot(aes(x = frac_cellt_IMC_MacCD163,y = frac_cellt_mIF_MacCD163))+
  geom_point()+
  geom_abline(slope = 1,intercept = 0)+
  coord_cartesian(xlim = c(0,1),ylim = c(0,1))+
  theme_bw()
cor(dat$frac_cellt_mIF_MacCD163,dat$frac_cellt_IMC_MacCD163)


```

# A - Number of cells steinbock

```{r,fig.height=15}
if (!dir.exists("output/Figure_S5/")) dir.create("output/Figure_S5/")

no_IMC <- colData(sce_IMC) %>% as_tibble() %>%
  group_by(sample_id, patient_id) %>%
  summarize(count = n())
# filter small cells
mif_steinbock <- mif_steinbock[,mif_steinbock$area > 28]

no_mIF <- colData(sce_mIF) %>% as_tibble() %>%
  group_by(sample_id, patient_id) %>%
  summarize(count = n())

no_mIF_steinbock <- colData(mif_steinbock) %>% as_tibble() %>%
  group_by(sample_id, patient_id) %>%
  summarize(count = n())

combined_df <- left_join(no_IMC, no_mIF, by = c("sample_id", "patient_id"))
combined_df <- left_join(combined_df, no_mIF_steinbock, by = c("sample_id", "patient_id"))
combined_df$patient_id <- sub("-", "", str_extract(combined_df$sample_id, "[0-9]{8}-"))

# difference steinbock vs IFQuant
combined_df %>%
  mutate(diff_size = count/(count.y/100)) %>%
  ungroup() %>%
  mutate(mean_size_dif = mean(diff_size))

p1 <- ggplot(combined_df) + 
  geom_point(aes(count.x, count.y, color = patient_id), size = 2) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red") +
  geom_smooth(aes(count.x, count.y), method='lm', formula= y~x, fullrange=TRUE) + 
  theme_minimal(base_size = 20) + xlab("# Cells in IMC") + ylab("# Cells in mIF") +
  coord_fixed() + 
  ylim(c(0, 5000)) + xlim(c(0, 5000)) +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$patient_id, name = "Patient ID")

p2 <- ggplot(combined_df) + 
  geom_point(aes(count.x, count, color = patient_id), size = 2) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red") +
  geom_smooth(aes(count.x, count), method='lm', formula= y~x, fullrange=TRUE) + 
  theme_minimal(base_size = 20) + xlab("# Cells in IMC") + ylab("# Cells in mIF steinbock") +
  coord_fixed() + 
  ylim(c(0, 5000)) + xlim(c(0, 5000)) +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$patient_id, name = "Patient ID")

p1
p2

ggsave(filename = "IMC_mIF_celltype_count.pdf",path = output, plot = p1, width = 10, height = 7)
ggsave(filename = "IMC_mIF-steinbock_celltype_count.pdf",path = output, plot = p2, width = 10, height = 7)
```

# check segmentation mIF
```{r, fig.height=10, fig.width=10}
mif_images <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/mIF/Rout/images.rds")
cur_im <- cytomapper::getImages(mif_images,i = 2)
cur_sce <- sce_mIF[,sce_mIF$sample_id == "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"]

# Load necessary library
library(tiff)

# Define image dimensions
width <- 1198
height <- 1196

# Create an empty matrix filled with zeros
image_matrix <- matrix(0, nrow = width, ncol = height)

# Example: Define x and y coordinates as a matrix
# Replace this with your actual x, y coordinate matrix
xy_coords <- as.matrix(data.frame(x = round(cur_sce$nucleus.x),
                        y = round(cur_sce$nucleus.y)))

# Assign running numbers to these specific pixels
for (i in 1:nrow(xy_coords)) {
  x <- xy_coords[i, 1]
  y <- xy_coords[i, 2]
  
  # Ensure coordinates are within bounds
    image_matrix[x, y] <- 1  # Running number
}

cur_mask <- EBImage::as.Image(image_matrix)
cur_mask <- cytomapper::CytoImageList(cur_mask)
mcols(cur_im)[,"sample_id"] <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"
names(cur_mask) <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"
mcols(cur_mask)$sample_id <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"

cytomapper::plotPixels(mask = cur_mask,
                       image = cur_im,
                       img_id = "sample_id",
                       colour_by = c("DAPI", "CD15"),
                       bcg = list(DAPI = c(0,1.5,1),
                                  CD15 = c(0,1,1)),save_plot = list(filename = "D:/Data/test_mIF.png"))

```

# check segmentation IMC
```{r}
IMC_images <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/IMC/Rout/images.rds")
cur_im <- cytomapper::getImages(IMC_images,i = 2)
cur_sce <- sce_IMC[,sce_IMC$sample_id == "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"]


# Load necessary library
library(tiff)

# Define image dimensions
width <- 600
height <- 599

# Create an empty matrix filled with zeros
image_matrix <- matrix(0, nrow = width, ncol = height)

# Example: Define x and y coordinates as a matrix
# Replace this with your actual x, y coordinate matrix
xy_coords <- as.matrix(data.frame(x = round(cur_sce$Pos_X),
                        y = round(cur_sce$Pos_Y)))

# Assign running numbers to these specific pixels
for (i in 1:nrow(xy_coords)) {
  x <- xy_coords[i, 1]
  y <- xy_coords[i, 2]
  
  # Ensure coordinates are within bounds
    image_matrix[x, y] <- 1  # Running number
}


cur_mask <- EBImage::as.Image(image_matrix)
cur_mask <- cytomapper::CytoImageList(cur_mask)
mcols(cur_im)[,"sample_id"] <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"
names(cur_mask) <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"
mcols(cur_mask)$sample_id <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"
cytomapper::plotPixels(mask = cur_mask, image = cur_im,img_id = "sample_id",colour_by = c("CD15", "DNA1", "CD3"),bcg = list(CD15 = c(0,2,1),
                                                                                                                            DAPI = c(0,2,1),CD3 = c(0,2,1)))

cytomapper::plotPixels(mask = cur_mask,
                       image = cur_im,
                       img_id = "sample_id",
                       colour_by = c("DNA2", "CD15"),
                       bcg = list(DNA2 = c(0,2.5,1),
                                  CD15 = c(0,1.5,1)),save_plot = list(filename = "D:/Data/test_IMC.png"))
```

# check segmentation mIF steinbock
```{r}
mif_st_images <- readRDS("O:/projects/immucan/processed_data/Panel_1/2022_WORKFLOW/mIF/old/steinbock/Rout/images.rds")
cur_im <- cytomapper::getImages(mif_st_images,i = 2)
cur_sce <- mif_steinbock[,mif_steinbock$sample_id == "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"]


# Load necessary library
library(tiff)

# Define image dimensions
width <- 1198
height <- 1196

# Create an empty matrix filled with zeros
image_matrix <- matrix(0, nrow = width, ncol = height)

# Example: Define x and y coordinates as a matrix
# Replace this with your actual x, y coordinate matrix
xy_coords <- as.matrix(data.frame(x = round(cur_sce$Pos_X),
                        y = round(cur_sce$Pos_Y)))

# Assign running numbers to these specific pixels
for (i in 1:nrow(xy_coords)) {
  x <- xy_coords[i, 1]
  y <- xy_coords[i, 2]
  
  # Ensure coordinates are within bounds
    image_matrix[x, y] <- 1  # Running number
}


cur_mask <- EBImage::as.Image(image_matrix)
cur_mask <- cytomapper::CytoImageList(cur_mask)
mcols(cur_im)[,"sample_id"] <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"
names(cur_mask) <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"
mcols(cur_mask)$sample_id <- "IMMUcan_2022_WFLOW_10061074-SPECT-VAR-TIS-01-IMC-01_002"
cytomapper::plotPixels(mask = cur_mask, image = cur_im,img_id = "sample_id",colour_by = c("CD15", "DAPI", "CD3"),bcg = list(CD15 = c(0,2,1),
                                                                                                                            DAPI = c(0,2,1),CD3 = c(0,2,1)))

cytomapper::plotPixels(mask = cur_mask,
                       image = cur_im,
                       img_id = "sample_id",
                       colour_by = c("DAPI", "CD15"),
                       bcg = list(DAPI = c(0,0.8,1),
                                  CD15 = c(0,1,1)),save_plot = list(filename = "D:/Data/test_mIF_steinbock.png"))
```

# compare numbers of cells
```{r}
imc_dat <- as.data.frame(colData(sce_IMC))
mif_dat <- as.data.frame(colData(sce_mIF))

imc_dat <- imc_dat %>%
  group_by(sample_id) %>%
  count(matched_celltype) %>%
  mutate(tech = "IMC")

mif_dat <- mif_dat %>%
  group_by(sample_id) %>%
  count(matched_celltype) %>%
  mutate(tech = "mif")

dat <- left_join(imc_dat,mif_dat,c("sample_id","matched_celltype"))
colnames(dat) <- c("sample_id","matched_celltype","imc_count","tech_imc","mif_count","tech_mif")

dat %>%
#  pivot_longer(cols = c(imc_count,mif_count)) %>%
  ggplot(aes(x = imc_count,y = mif_count,color = matched_celltype))+
  geom_point()+
  facet_wrap(.~matched_celltype, scales = "free")

(p_BnT <- ggplot(dat) +
  stat_smooth(aes(imc_count, mif_count, color = matched_celltype), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(imc_count, mif_count), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(imc_count, mif_count, color = matched_celltype), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") +
  ylim(c(0, 0.4)) + xlim(c(0, 0.4)) +
  stat_regline_equation(aes(imc_count, mif_count, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(imc_count, mif_count, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("BnT cells") +
  coord_fixed())
```

