---
title: "Figure S7"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: inline
---

Code to generate the plots of Figure S5.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Code to generate the plots of Figure 4.

# set path
the git repo contains a vector which specifies the path to the data folder. please adapt this path in your clone to run the repo automated.
```{r}
mount_path <- readRDS(file = "data/mount_path.rds")
output <- "output/Figure_S7"
```

## Read data

```{r, message=FALSE}
library(SingleCellExperiment)
library(tidyverse)
library(ggpubr)
library(cowplot)
library(imcRtools)


sce_IMC <- readRDS(file.path(mount_path, "IMC/Rout/sce.rds"))
sce_mIF <- readRDS(file.path(mount_path, "mIF/Rout/sce_ROIs.rds"))
colnames(sce_mIF) <- paste0(sce_mIF$sample_id,"_",sce_mIF$cell.ID)
```

# A - Number of cells

```{r}
if (!dir.exists("output/Figure_S7/")) dir.create("output/Figure_S7/")

no_IMC <- colData(sce_IMC) %>% as_tibble() %>%
  group_by(sample_id, patient_id) %>%
  summarize(count = n())
no_mIF <- colData(sce_mIF) %>% as_tibble() %>%
  group_by(sample_id, patient_id) %>%
  summarize(count = n())

combined_df <- left_join(no_IMC, no_mIF, by = c("sample_id", "patient_id"))
combined_df$patient_id <- sub("-", "", str_extract(combined_df$sample_id, "[0-9]{8}-"))

p <- ggplot(combined_df) + 
  geom_point(aes(count.x, count.y, color = patient_id), size = 2) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red") +
  geom_smooth(aes(count.x, count.y), method='lm', formula= y~x, fullrange=TRUE) + 
  theme_minimal(base_size = 20) + xlab("# Cells in IMC") + ylab("# Cells in mIF") +
  coord_fixed() + 
  ylim(c(0, 5000)) + xlim(c(0, 5000)) +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$patient_id, name = "Patient ID")

p
ggsave(filename = "FigS7_A.pdf",path = output, plot = p, width = 7, height = 5)
```

# B - cellular area

```{r}
IMC_dat <- data.frame(IMC_cell_size = sce_IMC$area)
mIF_dat <- data.frame(IF_cell_size = sce_mIF$cell.area)

mean_IMC <- mean(IMC_dat$IMC_cell_size)
mean_IF <- mean(mIF_dat$IF_cell_size )

cur_dat <- data.frame(IMC_cell_size = sce_IMC$area)

mean_IMC <- mean(cur_dat$IMC_cell_size)
mean_IF <- mean(sce_mIF$cell.area)

p <- cur_dat %>%
  ggplot(aes(y= IMC_cell_size))+
  geom_density(col= "blue", size = 2)+
  coord_flip()+
  theme_bw()+
  geom_density(data = mIF_dat ,aes(y = IF_cell_size),col= "red", size = 2)+
  annotate("text", x = 0.01, y = 392 , label = bquote("IMC: average" ~ .(round(mean_IMC)) ~  µm^2), color = "blue", size = 6)+
  annotate("text", x = 0.011, y = 400 , label =  bquote("mIF: average" ~ .(round(mean_IF)) ~  µm^2), color = "red", size = 6)+
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))+
  ylab("Cell size")+
  xlab("Density estimate")

p

ggsave(filename = "FigS7_c.pdf",path= output, plot = p, width = 7, height = 5)
```




# Supporting plots for DA analysis

```{r, fig.height=10, fig.width=10}
cur_mIF <- colData(sce_mIF)
cur_IMC <- colData(sce_IMC)

# Proportions of cell types for IMC
cur_IMC$matched_celltype <- factor(cur_IMC$matched_celltype)
prop_IMC <- cur_IMC %>% 
  as.data.frame() %>%
  group_by(sample_id, patient_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  group_by(sample_id) %>%
  mutate(prop = count/sum(count))

# Proportions of cell types for mIF
cur_mIF$matched_celltype <- factor(cur_mIF$matched_celltype)
prop_mIF <- cur_mIF %>% 
  as.data.frame() %>%
  group_by(sample_id, patient_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  group_by(sample_id) %>%
  mutate(prop = count/sum(count))

prop_IMC$modality <- "IMC"
prop_mIF$modality <- "mIF"

out <- rbind(prop_IMC, prop_mIF)

(p <- ggplot(out) + 
  geom_boxplot(aes(modality, prop), outlier.shape = NA) +
  geom_point(aes(modality, prop, color = patient_id)) +
  geom_line(aes(modality, prop, group = sample_id)) +
  facet_wrap(~matched_celltype, nrow = 1, scales = "free")  +
  theme_classic(base_size = 20) +
  ylab("Proportion of cell phenotypes") +
  xlab("Modality") +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$patient_id, "Patient ID"))
p
ggsave(filename = "FigS7_D.pdf",path=output, plot = p, height = 5, width = 17)
```

```{r}
p1 <- plotSpatial(sce_IMC[,sce_IMC$sample_id %in% c("IMMUcan_2022_WFLOW_10068868-SPECT-VAR-TIS-01-IMC-01_005")], 
            node_color_by = "matched_celltype", 
            img_id = "sample_id",
            node_size_fix = 1,) + 
  theme(strip.text =  element_text(size=1))+
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype)+
  ggtitle("IMC sample 10068868 ROI 005")+
  theme(legend.position = "bottom")
p1
p2 <- plotSpatial(sce_IMC[,sce_IMC$sample_id %in% c("IMMUcan_2022_WFLOW_10071582-SPECT-VAR-TIS-01-IMC-01_007")], 
            node_color_by = "matched_celltype", 
            img_id = "sample_id",
            node_size_fix = 1,) + 
  theme(strip.text =  element_text(size=1))+
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype)+
  ggtitle("IMC sample 10071582 ROI 007")+
  theme(legend.position = "bottom")


p3 <- plotSpatial(sce_mIF[,sce_mIF$sample_id %in% c("IMMUcan_2022_WFLOW_10068868-SPECT-VAR-TIS-01-IMC-01_005")],
                     coords = c("nucleus.x","nucleus.y"),
                     node_color_by = "matched_celltype", 
                     img_id = "sample_id",
                     node_size_fix = 1,) + 
  theme(strip.text =  element_text(size=1))+
  scale_color_manual(values = metadata(sce_mIF)$color_vectors$matched_celltype)+
  ggtitle("mIF sample 10068868 ROI 005")+
  theme(legend.position = "bottom")

p4 <- plotSpatial(sce_mIF[,sce_mIF$sample_id %in% c("IMMUcan_2022_WFLOW_10071582-SPECT-VAR-TIS-01-IMC-01_007")],
                     coords = c("nucleus.x","nucleus.y"),
                     node_color_by = "matched_celltype", 
                     img_id = "sample_id",
                     node_size_fix = 1,) + 
  theme(strip.text =  element_text(size=1))+
  scale_color_manual(values = metadata(sce_mIF)$color_vectors$matched_celltype)+
  ggtitle("mIF sample 10071582 ROI 007")

final_plot <- (p1 + p3 + p2 + p4) + 
  patchwork::plot_layout(guides = "collect",ncol = 4) &  # Collect the legends
  theme(legend.position = "bottom")
final_plot
ggsave(filename = "FigS7_E.pdf",path=output, plot = final_plot, height = 5, width = 15)
```


--------------------------------
The code below is code that is not part of the publication but some checks that were performed throughout
--------------------------------


Example images

```{r, eval=FALSE}
library(cytomapper)

(p1 <- plotSpatial(sce_IMC[,sce_IMC$sample_id == "IMMUcan_2022_WFLOW_10073140-SPECT-VAR-TIS-01-IMC-01_004"], 
            img_id = "sample_id", node_color_by = "matched_celltype") +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype) +
    theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA),
        legend.position = "none") + ggtitle(""))

(p2 <- plotSpatial(sce_mIF[,sce_mIF$sample_id == "IMMUcan_2022_WFLOW_10073140-SPECT-VAR-TIS-01-IMC-01_004"], 
            img_id = "sample_id", node_color_by = "matched_celltype", coords = c("nucleus.x", "nucleus.y")) +
  scale_color_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype) +
    theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_rect(fill = "black", colour = NA)) + ggtitle(""))

#ggsave(filename = "../output/Figure_S6/FigS5_D1.pdf", plot = p1, height = 10, width = 10)
#ggsave(filename = "../output/Figure_S6/FigS5_D2.pdf", plot = p2, height = 10, width = 10)
```


# whole slide mIF IMC correlation
we will correlate the fractions of cell types from mIF and IMC. we will therefore exclude ROI4 (TLS) which will otherwise falsify the results because they add many B cells to the IMC for a specific patient
```{r, fig.height=10, fig.width=10, eval=FALSE}
mIF_sce <- readRDS(paste0(mount_path,"mIF/Rout/sce_whole_slide.rds"))

unique(mIF_sce$patient_id)
unique(sce_IMC$patient_id)

cur_imc <- sce_IMC[,sce_IMC$ROI != 4]

mIF <- as.data.frame(colData(mIF_sce))
IMC <- as.data.frame(colData(sce_IMC))
mIF <- mIF %>%
  group_by(patient_id) %>%
  count(matched_celltype, name = "ct_count_mIF") %>%
  mutate(total_cell_count_mIF = sum(ct_count_mIF),
         frac_cellt_mIF = ct_count_mIF/total_cell_count_mIF)

IMC <- IMC %>%
  filter(! ROI %in% c(4,8)) %>%
  group_by(patient_id) %>%
  count(matched_celltype, name = "ct_count_IMC") %>%
  mutate(total_cell_count_IMC = sum(ct_count_IMC),
         frac_cellt_IMC = ct_count_IMC/total_cell_count_IMC)

dat <- left_join(IMC, mIF, c("patient_id","matched_celltype"))

dat <- dat %>%
  pivot_wider(id_cols = patient_id,names_from = matched_celltype,values_from = c(frac_cellt_IMC,frac_cellt_mIF))
cor <- psych::corr.test(dat[,grepl("_IMC",colnames(dat))],dat[,grepl("mIF",colnames(dat))],  method = "pearson")

cur_dat <- as.data.frame(cor$r)
cur_dat$variable <- rownames(cur_dat)
dat_long <- reshape2::melt(cur_dat,id.vars="variable")
colnames(dat_long) <- c("cell_types_IMC","cell_types_mIF","correlation")

#get p-value matrix
p_dat <- as_tibble(cor$p)
p_dat$variable <- rownames(cur_dat)
pdat_long <- reshape2::melt(p_dat,id.vars="variable")
colnames(pdat_long) <- c("cell_types_IMC","cell_types_mIF","p_adj")

dat_long$p_adj <- pdat_long$p_adj

dat_long <- dat_long %>%
  mutate(sig = ifelse(p_adj <= 0.001 & p_adj > 0.0001,0.001,p_adj))

dat_long <- dat_long %>%
  mutate(sig = case_when(p_adj <= 0.0001 ~ "< 0.0001",
                         p_adj <= 0.001 & p_adj > 0.0001 ~ "< 0.001",
                         p_adj <= 0.01 & p_adj > 0.001 ~ "< 0.01",
                         p_adj <= 0.1 & p_adj > 0.01 ~ "< 0.1",
                         p_adj >0.1 ~ "ns"))

cor_filter <- 0.0
library(ggnewscale)
p <- dat_long %>%
   filter(abs(correlation) > cor_filter) %>%
   ggplot(aes(x=factor(cell_types_IMC, levels=sort(unique(dat_long$cell_types_IMC))),y = factor(cell_types_mIF, levels=sort(levels(dat_long$cell_types_mIF))),col = correlation, size= -log10(p_adj)))+
    #geom_tile(aes(fill = sig))+
    #theme_bw()+
    geom_point()+
    theme_bw()+
    scale_color_gradient2(low =  "blue",mid = "white",high = "red", name = expression("Pearson's" ~ rho))+
    #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    #      panel.background = element_blank(), axis.line = element_line(colour = "black"))+
    #theme_bw()+
    #scale_color_gradient(low = "white",   high = "red")+
    scale_size(range = c(1,8))+
    new_scale_colour() +
    geom_point(data = dat_long %>% filter(sig <0.01, abs(correlation) > cor_filter),shape = 1,aes(size = -log10(p_adj), col = sig), stroke =2.5)+
    scale_color_manual(values = c("TRUE" = "black","FALSE" = "white"), name = "FDR < 0.1")+
    theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1,size = 20))+
    theme(axis.text.y = element_text(size = 20))+
    theme(legend.text = element_text(size = 20))+
    theme(legend.title = element_text(size = 20))
p

ggsave(filename = paste0(output,"/wholeslide_mIF_IMC_correlation.pdf"),plot = p,width = 12, height = 8)
```

# matched region mIF IMC correlation
```{r, fig.height=10, fig.width=10, eval=FALSE}
unique(mIF_sce$patient_id)
unique(sce_IMC$patient_id)

cur_imc <- sce_IMC[,sce_IMC$ROI != 4]

mIF <- as.data.frame(colData(sce_mIF))
IMC <- as.data.frame(colData(sce_IMC))
mIF <- mIF %>%
  group_by(patient_id) %>%
  count(matched_celltype, name = "ct_count_mIF") %>%
  mutate(total_cell_count_mIF = sum(ct_count_mIF),
         frac_cellt_mIF = ct_count_mIF/total_cell_count_mIF)

IMC <- IMC %>%
  filter(! ROI %in% c(4,8)) %>%
  group_by(patient_id) %>%
  count(matched_celltype, name = "ct_count_IMC") %>%
  mutate(total_cell_count_IMC = sum(ct_count_IMC),
         frac_cellt_IMC = ct_count_IMC/total_cell_count_IMC)

dat <- left_join(IMC, mIF, c("patient_id","matched_celltype"))

dat <- dat %>%
  pivot_wider(id_cols = patient_id,names_from = matched_celltype,values_from = c(frac_cellt_IMC,frac_cellt_mIF))
cor <- psych::corr.test(dat[,grepl("_IMC",colnames(dat))],dat[,grepl("mIF",colnames(dat))],  method = "pearson")

cur_dat <- as.data.frame(cor$r)
cur_dat$variable <- rownames(cur_dat)
dat_long <- reshape2::melt(cur_dat,id.vars="variable")
colnames(dat_long) <- c("cell_types_IMC","cell_types_mIF","correlation")

#get p-value matrix
p_dat <- as_tibble(cor$p)
p_dat$variable <- rownames(cur_dat)
pdat_long <- reshape2::melt(p_dat,id.vars="variable")
colnames(pdat_long) <- c("cell_types_IMC","cell_types_mIF","p_adj")

dat_long$p_adj <- pdat_long$p_adj

dat_long <- dat_long %>%
  mutate(sig = ifelse(p_adj <= 0.001 & p_adj > 0.0001,0.001,p_adj))

dat_long <- dat_long %>%
  mutate(sig = case_when(p_adj <= 0.0001 ~ "< 0.0001",
                         p_adj <= 0.001 & p_adj > 0.0001 ~ "< 0.001",
                         p_adj <= 0.01 & p_adj > 0.001 ~ "< 0.01",
                         p_adj <= 0.1 & p_adj > 0.01 ~ "< 0.1",
                         p_adj >0.1 ~ "ns"))

cor_filter <- 0.0
p <- dat_long %>%
   filter(abs(correlation) > cor_filter) %>%
   ggplot(aes(x=factor(cell_types_IMC, levels=sort(unique(dat_long$cell_types_IMC))),y = factor(cell_types_mIF, levels=sort(levels(dat_long$cell_types_mIF))),col = correlation, size= -log10(p_adj)))+
    #geom_tile(aes(fill = sig))+
    #theme_bw()+
    geom_point()+
    theme_bw()+
    scale_color_gradient2(low =  "blue",mid = "white",high = "red", name = expression("Pearson's" ~ rho))+
    #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    #      panel.background = element_blank(), axis.line = element_line(colour = "black"))+
    #theme_bw()+
    #scale_color_gradient(low = "white",   high = "red")+
    scale_size(range = c(1,8))+
    new_scale_colour() +
    geom_point(data = dat_long %>% filter(sig <0.01, abs(correlation) > cor_filter),shape = 1,aes(size = -log10(p_adj), col = sig), stroke =2.5)+
    scale_color_manual(values = c("TRUE" = "black","FALSE" = "white"), name = "FDR < 0.1")+
    theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1,size = 20))+
    theme(axis.text.y = element_text(size = 20))+
    theme(legend.text = element_text(size = 20))+
    theme(legend.title = element_text(size = 20))
p

ggsave(filename = paste0(output,"/matched_mIF_IMC_correlation.pdf"),plot = p,width = 12, height = 8)
```



# steinbock comparison
```{r, fig.width=10, fig.height=8, eval=FALSE}
sce_st <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/steinbock_2025/Rout/sce.rds")
sce_nuc <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/steinbock_2025/Rout/sce_nuc.rds")

st_dat <- as.data.frame(t(assay(sce_st,"counts")))
st_nuc <- as.data.frame(t(assay(sce_nuc,"counts")))
mif_dat <- as.data.frame(t(assay(sce_mIF,"exprs")))

LSD::heatscatter(log(st_dat$CD3), log(st_dat$CD11c))
LSD::heatscatter(log(st_nuc$CD3), log(st_nuc$CD11c))
LSD::heatscatter(mif_dat$CD3, mif_dat$CD11c)

images_IMC <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/IMC/Rout/images.rds")
masks_IMC <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/IMC/Rout/masks.rds")
images_mIF <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/mIF/Rout/images.rds")
images_st <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/steinbock_2025/Rout/images.rds")
masks_st <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/steinbock_2025/Rout/masks.rds")
# check specific samples
 unique(sce_IMC$sample_id)[grepl("10074349", unique(sce_IMC$sample_id))]
for (i in unique(sce_IMC$sample_id)[grepl("10074349", unique(sce_IMC$sample_id))]) {
  
  imc_dat <- as.data.frame(t(assay(sce_IMC[,sce_IMC$sample_id == i],"exprs")))
  st_dat <- as.data.frame(t(assay(sce_st[,sce_st$sample_id == i],"counts")))
  st_nuc <- as.data.frame(t(assay(sce_nuc[,sce_nuc$sample_id == i],"counts")))
  mif_dat <- as.data.frame(t(assay(sce_mIF[,sce_mIF$sample_id == i],"exprs")))
  
  # IMC images with cells
  cytomapper::plotPixels(sce_IMC[,sce_IMC$sample_id == i & sce_IMC$celltype == "DC"],
                         cell_id = "ObjectNumber",
                         mask = cytomapper::getImages(masks_IMC,i = i),
                         image = cytomapper::getImages(images_IMC,i = i),
                         img_id = "sample_id",
                         outline_by = "matched_celltype",
                         colour_by = c("DNA1", "CD11c", "CD3"),
                         bcg = list(DNA1 = c(0,2,1),
                                    CD11c = c(0,3,1),
                                    CD3= c(0,5,1)))
  # mIF images without cells
  cytomapper::plotPixels(image = cytomapper::getImages(images_mIF,i = i),
                         img_id = "sample_id",
                         colour_by = c("DAPI", "CD11c", "CD3"),
                         bcg = list(DAPI = c(0,1,1),
                                    CD11c = c(0,1,2),
                                    CD3 = c(0,5,1)))
  dittoSeq::dittoScatterPlot(sce_IMC[,sce_IMC$sample_id == i],x.var = "CD163",y.var = "CD11c",assay.x = "exprs", assay.y = "exprs",color.var = "matched_celltype")
  dittoSeq::dittoScatterPlot(sce_mIF2[,sce_mIF2$sample_id == i],x.var = "CD163",y.var = "CD11c",assay.x = "exprs", assay.y = "exprs",color.var = "matched_celltype")
  dittoSeq::dittoScatterPlot(sce_IMC[,sce_IMC$sample_id == i],x.var = "CD3",y.var = "CD11c",assay.x = "exprs", assay.y = "exprs",color.var = "matched_celltype")
  dittoSeq::dittoScatterPlot(sce_mIF2[,sce_mIF2$sample_id == i],x.var = "CD3",y.var = "CD11c",assay.x = "exprs", assay.y = "exprs",color.var = "matched_celltype")
  dittoSeq::dittoScatterPlot(sce_nuc[,sce_nuc$sample_id == i],x.var = "CD3",y.var = "CD11c",assay.x = "exprs", assay.y = "exprs")
  
  
  plotSpatial(sce_mIF[,sce_mIF$sample_id == i],img_id = "sample_id",coords = c("nucleus.x","nucleus.y"),node_color_by = "matched_celltype")+scale_colour_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype)
  
  plotSpatial(sce_IMC[,sce_IMC$sample_id == i],img_id = "sample_id",node_color_by = "matched_celltype")+scale_colour_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype)
  
  
  
  plotSpatial(sce_st[,sce_st$sample_id == i],img_id = "sample_id",node_color_by = "CD11c", assay_type = "exprs")
  plotSpatial(sce_nuc[,sce_nuc$sample_id == i],img_id = "sample_id",node_color_by = "CD11c", assay_type = "exprs")
  plotSpatial(sce_mIF[,sce_mIF$sample_id == i],img_id = "sample_id",coords = c("nucleus.x","nucleus.y"),node_color_by = "CD11c", assay_type = "exprs")
  plotSpatial(sce_IMC[,sce_IMC$sample_id == i],img_id = "sample_id",node_color_by = "CD11c", assay_type = "exprs")
}

sce_mIF$matched_celltype

```

# test if we alter the phenotype key how that changes DC correlation
```{r, eval=FALSE}
cur_key <- as.data.frame(read_csv(paste0(mount_path,"mIF/phenotype_key_IF1.csv")))
cur_key <- cur_key %>%
  mutate(celltype = ifelse(phenotype == "CD15-CK-CD3+CD11c+CD20+CD163-","DC",celltype))
rownames(cur_key) <- cur_key$phenotype
sce_mIF2 <- sce_mIF
sce_mIF2$celltype <- cur_key[sce_mIF$phenotype,2]
sce_mIF2$matched_celltype <- sce_mIF2$celltype
sce_mIF2$matched_celltype[grepl("Tumor", sce_mIF2$matched_celltype)] <- "Tumor"
```

## plot altered phenotype key
```{r, eval=FALSE}
# Match cell types
cur_mIF <- colData(sce_mIF2)
cur_IMC <- colData(sce_IMC)

cur_mIF$sample_id <- as.factor(cur_mIF$sample_id)
cur_mIF$matched_celltype <- as.factor(cur_mIF$matched_celltype)
cur_IMC$sample_id <- as.factor(cur_IMC$sample_id)
cur_IMC$matched_celltype <- as.factor(cur_IMC$matched_celltype)

frac_IMC <- cur_IMC %>% as_tibble() %>%
  group_by(sample_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

frac_mIF <- cur_mIF %>% as_tibble() %>%
  group_by(sample_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

combined <- left_join(frac_IMC, frac_mIF, by=c("sample_id", "matched_celltype"))
combined$patient_id <- sub("-", "", str_extract(combined$sample_id, "[0-9]{8}-"))

(p_B <- ggplot(combined[combined$matched_celltype == "B",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.3)) + xlim(c(0, 0.3)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("B cells") +
  coord_fixed())

(p_BnT <- ggplot(combined[combined$matched_celltype == "BnT",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.4)) + xlim(c(0, 0.4)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("BnT cells") +
  coord_fixed())

(p_T <- ggplot(combined[combined$matched_celltype == "T",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.6)) + xlim(c(0, 0.6)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("T cells") +
  coord_fixed())

(p_Mac <- ggplot(combined[combined$matched_celltype == "MacCD163",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.3)) + xlim(c(0, 0.3)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Macrophages") +
  coord_fixed())

(p_DC_new <- ggplot(combined[combined$matched_celltype == "DC",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.2)) + xlim(c(0, 0.2)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("DC_new_key") +
  coord_fixed())

(p_Neutro <- ggplot(combined[combined$matched_celltype == "Neutrophil",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.2)) + xlim(c(0, 0.2)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Neutrophils") +
  coord_fixed())

(p_Tumor <- ggplot(combined[combined$matched_celltype == "Tumor",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.8)) + xlim(c(0, 0.8)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Tumor") +
  coord_fixed())

(p_other <- ggplot(combined[combined$matched_celltype == "other",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.7)) + xlim(c(0, 0.7)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Other cells") +
  coord_fixed())

```


```{r, eval=FALSE}
cur_mIF <- colData(sce_mIF)
cur_IMC <- colData(sce_IMC)

cur_mIF$sample_id <- as.factor(cur_mIF$sample_id)
cur_mIF$matched_celltype <- as.factor(cur_mIF$matched_celltype)
cur_IMC$sample_id <- as.factor(cur_IMC$sample_id)
cur_IMC$matched_celltype <- as.factor(cur_IMC$matched_celltype)

frac_IMC <- cur_IMC %>% as_tibble() %>%
  group_by(sample_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

frac_mIF <- cur_mIF %>% as_tibble() %>%
  group_by(sample_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

combined <- left_join(frac_IMC, frac_mIF, by=c("sample_id", "matched_celltype"))
combined$patient_id <- sub("-", "", str_extract(combined$sample_id, "[0-9]{8}-"))
(p_DC <- ggplot(combined[combined$matched_celltype == "DC",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.2)) + xlim(c(0, 0.2)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("DC_old_key") +
  coord_fixed())

p <- plot_grid(p_DC,p_DC_new)
p
ggsave(filename = paste0(output,"/DC_new_phenotype_key.png"),plot = p,device = "png",width = 3000, height = 1500,units = "px")
```



# Neutrophils
```{r, fig.width=15, fig.height=6, eval=FALSE}
pie(1:10, col = metadata(sce_IMC)$color_vectors$patient_id,labels = names(metadata(sce_IMC)$color_vectors$patient_id ))

pat <- "IMMUcan_2022_WFLOW_10068868-SPECT-VAR-TIS-01-IMC-01_008"
for (i in pat) {
  
  imc_dat <- as.data.frame(t(assay(sce_IMC[,sce_IMC$sample_id == i],"exprs")))
  st_dat <- as.data.frame(t(assay(sce_st[,sce_st$sample_id == i],"counts")))
  st_nuc <- as.data.frame(t(assay(sce_nuc[,sce_nuc$sample_id == i],"counts")))
  mif_dat <- as.data.frame(t(assay(sce_mIF[,sce_mIF$sample_id == i],"exprs")))
  
  # IMC images with cells
  cytomapper::plotPixels(sce_IMC[,sce_IMC$sample_id == i & sce_IMC$celltype == "Neutrophil"],
                         cell_id = "ObjectNumber",
                         mask = cytomapper::getImages(masks_IMC,i = i),
                         image = cytomapper::getImages(images_IMC,i = i),
                         img_id = "sample_id",
                         outline_by = "matched_celltype",
                         colour_by = c("DNA1", "Ecad", "CD15"),
                         bcg = list(DNA1 = c(0,2,1),
                                    Ecad = c(0,5,1),
                                    CD15= c(0,5,1)))
  # mIF images without cells
  cytomapper::plotPixels(image = cytomapper::getImages(images_mIF,i = i),
                         img_id = "sample_id",
                         colour_by = c("DAPI", "CK", "CD15"),
                         bcg = list(DAPI = c(0,1,1),
                                    CK = c(0,3,1),
                                    CD15 = c(0,5,1)))
  pIMC <- dittoSeq::dittoScatterPlot(sce_IMC[,sce_IMC$sample_id == i],x.var = "MPO",y.var = "Ecad",assay.x = "exprs", assay.y = "exprs",color.var = "matched_celltype")
  pmif <- dittoSeq::dittoScatterPlot(sce_mIF[,sce_mIF$sample_id == i],x.var = "CD15",y.var = "CK",assay.x = "exprs", assay.y = "exprs",color.var = "matched_celltype")
  print(pIMC+ ggtitle("IMC"))
  print(pmif+ggtitle("mIF"))
}

pIMC <- pIMC+ ggtitle(paste0(i, " - ","IMC"))+scale_color_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype)+
  theme(plot.title = element_text(size = 7, face = "bold"))
pmif <- pmif+ggtitle(paste0(i, " - ","mIF"))+scale_color_manual(values = metadata(sce_IMC)$color_vectors$matched_celltype)+
  theme(plot.title = element_text(size = 7, face = "bold"))
p<- plot_grid(pIMC, pmif)
p
ggsave(filename = paste0(output,"/neutrophils_IMC_mif.png"),plot = p,device = "png",width = 3000, height = 1500,units = "px")

```

# gaussian mixture to identify phenotypes
```{r, fig.height=4,fig.width=8, eval=FALSE}
sce_nuc <- readRDS("D:/Data/2022_Workflow_for_zenodo_upload/steinbock_2025/Rout/sce.rds")
colnames(sce_nuc) <- paste0(sce_nuc$sample_id,"_",sce_nuc$ObjectNumber)

library(mclust)
sce_nuc$phenotype <- ""

p.list <- list()
for(i in unique(sce_nuc$sample_id)){
  cur_sce <- sce_nuc[,sce_nuc$sample_id == i]
  cur_dat <- as.data.frame(t(assay(cur_sce,"exprs")))
  for (k in c("CD15","CK","CD3","CD11c","CD20","CD163")) {
    
    fit <- Mclust(cur_dat[,k],G=2,modelNames = "E")
    cur_sce$phenotype <- fit$classification
    p1 <- dittoSeq::dittoScatterPlot(cur_sce[,cur_sce$sample_id == i],x.var = k,y.var = "CD3", assay.x = "exprs", assay.y = "exprs", color.var = "phenotype")+
      ggtitle(paste0(i,"__",k))
    cur_out <- data.frame(value = t(assay(cur_sce,"exprs"))[,k],
                          phenotype = cur_sce$phenotype,
                          variable = k)
    p2 <- cur_out %>%
      ggplot()+
      geom_density(aes(x = value,fill = as.factor(phenotype)))+
      ggtitle(paste0(i,"__",k))
    
    p.list[[paste0(i,"__",k)]] <- cowplot::plot_grid(p1,p2)
    cur_sce$phenotype <- ifelse(cur_sce$phenotype == 1,paste0(k,"-"),paste0(k,"+"))
    p_vec <- cur_sce$phenotype
    names(p_vec) <- colnames(cur_sce)
    sce_nuc[,names(p_vec)]$phenotype <- paste0(sce_nuc[,names(p_vec)]$phenotype,p_vec[colnames(sce_nuc[,sce_nuc$sample_id == i])])
  }
}
```

# apply phenotype key
```{r, eval=FALSE}
cur_key <- as.data.frame(read_csv(paste0(mount_path,"mIF/phenotype_key_IF1.csv")))
rownames(cur_key) <- cur_key$phenotype

sce_nuc$celltype <- cur_key[sce_nuc$phenotype,2]
sce_nuc$matched_celltype <- sce_nuc$celltype
sce_nuc$matched_celltype[grepl("Tumor", sce_nuc$matched_celltype)] <- "Tumor"
```

## Cell phenotype correlation

```{r, eval=FALSE}
# Match cell types
cur_mIF <- colData(sce_nuc)
cur_IMC <- colData(sce_IMC)

cur_mIF$sample_id <- as.factor(cur_mIF$sample_id)
cur_mIF$matched_celltype <- as.factor(cur_mIF$matched_celltype)
cur_IMC$sample_id <- as.factor(cur_IMC$sample_id)
cur_IMC$matched_celltype <- as.factor(cur_IMC$matched_celltype)

frac_IMC <- cur_IMC %>% as_tibble() %>%
  group_by(sample_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

frac_mIF <- cur_mIF %>% as_tibble() %>%
  group_by(sample_id, matched_celltype, .drop = FALSE) %>%
  summarize(count = n()) %>%
  mutate(frac = count/sum(count))

combined <- left_join(frac_IMC, frac_mIF, by=c("sample_id", "matched_celltype"))
combined$patient_id <- sub("-", "", str_extract(combined$sample_id, "[0-9]{8}-"))

(p_B <- ggplot(combined[combined$matched_celltype == "B",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.3)) + xlim(c(0, 0.3)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("B cells") +
  coord_fixed())

(p_BnT <- ggplot(combined[combined$matched_celltype == "BnT",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.4)) + xlim(c(0, 0.4)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("BnT cells") +
  coord_fixed())

(p_T <- ggplot(combined[combined$matched_celltype == "T",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.6)) + xlim(c(0, 0.6)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("T cells") +
  coord_fixed())

(p_Mac <- ggplot(combined[combined$matched_celltype == "MacCD163",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.3)) + xlim(c(0, 0.3)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Macrophages") +
  coord_fixed())

(p_DC <- ggplot(combined[combined$matched_celltype == "DC",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.2)) + xlim(c(0, 0.2)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("DC") +
  coord_fixed())

(p_Neutro <- ggplot(combined[combined$matched_celltype == "Neutrophil",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.2)) + xlim(c(0, 0.2)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Neutrophils") +
  coord_fixed())

(p_Tumor <- ggplot(combined[combined$matched_celltype == "Tumor",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.8)) + xlim(c(0, 0.8)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Tumor") +
  coord_fixed())

(p_other <- ggplot(combined[combined$matched_celltype == "other",]) +
  stat_smooth(aes(frac.x, frac.y, color = patient_id), alpha = 0.5, se = FALSE, method='lm', formula = y~x, fullrange=FALSE, geom = "line", lwd = 1.5) + 
  geom_smooth(aes(frac.x, frac.y), se = TRUE, method='lm', formula = y~x, fullrange=TRUE, lwd = 1.5) + 
  geom_point(aes(frac.x, frac.y, color = patient_id), size = 3) + 
  geom_abline(slope = 1, intercept = 0, color = "dark red", lwd = 1.5) +
  theme_classic(base_size = 20) + xlab("") + ylab("") + 
  scale_color_manual(values = c(metadata(sce_IMC)$color_vectors$ROIs, metadata(sce_IMC)$color_vectors$patient_id)) +
  ylim(c(0, 0.7)) + xlim(c(0, 0.7)) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..eq.label..), label.x.npc = 0, label.y.npc = 1, size = 10) +
  stat_regline_equation(aes(frac.x, frac.y, label = ..rr.label..), label.x.npc = 0, label.y.npc = 0.9, size = 10) +
  theme(legend.position = "none",
        axis.text = element_text(size = 20)) +
  ggtitle("Other cells") +
  coord_fixed())
```

```{r, fig.height=10, fig.width=20, eval=FALSE}
p <- plot_grid(p_T, p_B, p_BnT, p_Tumor, p_Mac, p_DC, p_other, p_Neutro, nrow = 2) 
p
ggsave(filename = paste0(output,"/mesmer_whole_cell_mIF_celltype_correlation.pdf"),device = "pdf",width = 20,height = 10)
```

